---
title: "Analysis & Key Findings"
subtitle: "Reconstructing the Prompt-Injection Worm at Tenant Thread"
date: "2026-06-10"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  message: false
  freeze: true
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
---

## 1. Overview

::: callout-important
#### Key Findings at a Glance

-   The **17 May 2046, 4:21 am** SaidIT post (challenge-stated local time; **11:21 UTC** in the log) was **not human-authored**. The log records `Agent/person:john_windward` receiving a task, passing a SaidIT permission check, posting `SwiftWren.txt`, then deleting both evidence files in a four-second burst.
-   The observed mechanism is a **prompt-injection worm**: an agent reads `*_further_instructions.md`, executes a file-sourced post, deletes the payload, and relays the same task to other agents.
-   The May 17 post was the **third occurrence**, following HiddenOrca on **10 May** and MellowOtter on **11 May**. This is a repeatable system failure, not an isolated glitch.
-   A **single orchestration-layer rule** - block `queue_subordinate_task` tasks whose path matches `*_further_instructions.md` - would have stopped all **235** malicious queue relays while leaving **16,803** legitimate queue events untouched.
:::

On **17 May 2046 at approximately 4:21 am challenge-stated local time** (**11:21 UTC**), a post appeared on SaidIT, Tenant Thread's internal social platform, attributed to John Windward (Department Lead, Customer Support). The content looked like gibberish because the post body was not typed into the post event - the log points to `content_source = "SwiftWren.txt"`. The system did exactly what a compromised agent instructed it to do: read a local payload file, publish it to `general`, and erase the artefacts.

This analysis uses the full 185,147-event timeline (`MC2 data.json`) and organisational chart (`org_chart.json`) to reconstruct the root cause, situate it within the broader system, trace historical recurrence, and propose a targeted, evidence-backed remediation.

### 1.1 Challenge Question Answer Map

```{r}
#| label: answer-map

suppressPackageStartupMessages(
  pacman::p_load(DT, dplyr, tibble)
)

answer_map <- tibble(
  `MC2 Question` = c(
    "1a. Exact chain for the May 17 post",
    "1b. System overview around that chain",
    "2. Meaning and origin of post contents",
    "3a. Prior occurrences and recurrence risk",
    "3b. One intervention point"
  ),
  `Direct Answer` = c(
    "Chloe Ballast's agent queued John Windward's agent to read SwiftWren_further_instructions.md; agent checked SaidIT access, posted SwiftWren.txt to 'general', then deleted both files within 4 seconds.",
    "The chain sits inside a 186-relay, 18-agent, 6-department relay network. Any SaidIT-capable agent was a viable terminal executor.",
    "Posts are file-sourced execution payloads, not normal messages. SwiftWren.txt traced to Emma Harbor (CFO); MellowOtter.txt to Noah Mariner (COO).",
    "Yes. HiddenOrca (May 10) and MellowOtter (May 11) preceded SwiftWren. Escalating scale and timed propagation suggest iterative worm refinement.",
    "Block queue_subordinate_task whose path ends in _further_instructions.md: 235/235 malicious relays blocked, 0/16,803 legitimate events affected."
  ),
  `Primary Evidence` = c(
    "Table 2 and Figure 6",
    "Figures 1, 5b, 7, 8, 8b, 9",
    "Table 3, Figure 10A, Figure 10, Table 4",
    "Figures 11, 12, 13, 13b and Table 5",
    "Figures 14, 15, 16 and Tables 6, 8"
  )
)

datatable(
  answer_map,
  options = list(dom = "t", paging = FALSE, scrollX = TRUE,
                 columnDefs = list(
                   list(width = "18%", targets = 0),
                   list(width = "60%", targets = 1),
                   list(width = "22%", targets = 2)
                 )),
  rownames = FALSE
)
```

**Analysis**

- All five challenge questions ultimately reduce to a single mechanical answer: one regex rule (`_further_instructions.md`) blocks every malicious event with zero false positives - a forensic result unusual in its precision.
- The chain of questions is deliberately ordered from the specific (what happened on May 17) to the systemic (why does this keep happening) - each section creates the context the next one needs to land.
- The unresolved item - HiddenOrca's origin - is the most critical open risk: two campaigns are closed; the entry channel for the first one is not, meaning the organisation has not definitively shut the door.

---

## 2. Data Preparation

### 2.1 Package Loading

```{r}
#| label: load-packages

pacman::p_load(
  jsonlite,      # JSON parsing
  tidyverse,     # data wrangling and ggplot2
  lubridate,     # datetime handling
  igraph,        # graph analytics and centrality
  visNetwork,    # interactive HTML networks
  echarts4r,     # interactive org chart tree
  plotly,        # interactive charts
  DT,            # interactive tables
  htmltools,     # combining HTML widgets and legends
  scales,        # axis formatting
  glue           # string interpolation
)

`%||%` <- function(x, y) if (is.null(x)) y else x
```

### 2.2 Data Loading

```{r}
#| label: load-data

mc2_raw     <- fromJSON("../VAST_Challenge_2026_MC2/MC2 data.json",
                        simplifyVector = FALSE)
events_list <- mc2_raw$events

events_df <- map_dfr(events_list, function(e) {
  tibble(
    id          = e$id         %||% NA_integer_,
    short_name  = e$short_name  %||% NA_character_,
    when        = e$when        %||% NA_real_,
    parties_str = paste(unlist(e$parties %||% list()), collapse = ";"),
    details_str = as.character(toJSON(e$details %||% list(), auto_unbox = TRUE))
  )
}) %>%
  mutate(
    datetime   = as_datetime(when),
    date       = as_date(datetime),
    hour       = hour(datetime),
    event_week = floor_date(datetime, "week"),
    event_day  = floor_date(datetime, "day")
  )

org_raw   <- fromJSON("../VAST_Challenge_2026_MC2/org_chart.json")
org_nodes <- as_tibble(org_raw$nodes)
org_edges <- as_tibble(org_raw$edges)
```

### 2.3 Feature Engineering

```{r}
#| label: feature-engineering

events_df <- events_df %>%
  mutate(
    is_anomalous_post = str_detect(details_str, "content_source"),
    actor_type = case_when(
      str_detect(parties_str, "^Agent/")  ~ "Agent",
      str_detect(parties_str, "^person:") ~ "Person",
      str_detect(parties_str, "^system:") ~ "System",
      TRUE                                ~ "Mixed"
    )
  )

parties_long <- events_df %>%
  select(id, short_name, datetime, date, details_str, parties_str) %>%
  mutate(parties = str_split(parties_str, ";")) %>%
  unnest(parties) %>%
  filter(parties != "") %>%
  mutate(
    party_type = case_when(
      str_starts(parties, "Agent/")  ~ "agent",
      str_starts(parties, "person:") ~ "person",
      str_starts(parties, "system:") ~ "system",
      TRUE                           ~ "other"
    ),
    person_id = str_remove(parties, "^Agent/")
  )

instr_pattern <- "_further_instructions\\.md"

chain_events <- events_df %>%
  filter(
    short_name %in% c("queue_subordinate_task", "assign_agent_task"),
    str_detect(details_str, instr_pattern)
  ) %>%
  mutate(
    instr_file = case_when(
      str_detect(details_str, "HiddenOrca")  ~ "HiddenOrca",
      str_detect(details_str, "MellowOtter") ~ "MellowOtter",
      str_detect(details_str, "SwiftWren")   ~ "SwiftWren"
    ),
    sender   = str_remove(
                 str_extract(parties_str, "Agent/person:[^;]+"), "^Agent/"),
    receiver = str_remove(
                 str_extract(details_str,  "Agent/person:[^\"]+"), "^Agent/")
  )

team_dept_lookup <- org_edges %>%
  filter(str_starts(source, "department:"), str_starts(target, "team:")) %>%
  left_join(org_nodes %>% select(id, dept_label = label),
            by = c("source" = "id")) %>%
  left_join(org_nodes %>% select(id, team_label = label),
            by = c("target" = "id")) %>%
  select(dept_id = source, team_id = target, dept_label, team_label)

team_members <- org_edges %>%
  filter(str_starts(source, "team:"), str_starts(target, "person:")) %>%
  left_join(team_dept_lookup, by = c("source" = "team_id")) %>%
  select(person_id = target, dept_label, team_label)

dept_members <- org_edges %>%
  filter(str_starts(source, "department:"), str_starts(target, "person:")) %>%
  left_join(org_nodes %>% select(id, dept_label = label),
            by = c("source" = "id")) %>%
  mutate(team_label = if_else(relation == "led_by",
                              paste0(dept_label, " Leadership"),
                              dept_label)) %>%
  select(person_id = target, dept_label, team_label)

person_org_lookup <- bind_rows(team_members, dept_members) %>%
  distinct(person_id, dept_label, team_label)

person_dept_lookup <- person_org_lookup %>%
  select(person_id, dept_label) %>%
  distinct()

clean_name <- function(x) {
  x %>%
    str_remove("^person:") %>%
    str_replace_all("_", " ") %>%
    str_to_title()
}

campaign_order <- c("HiddenOrca", "MellowOtter", "SwiftWren")

perm_surface <- events_df %>%
  filter(short_name == "saidit_post_check") %>%
  mutate(
    checker     = str_remove(str_extract(parties_str, "(Agent/)?person:[^;]+"), "^Agent/"),
    check_pass  = case_when(
      str_detect(details_str, '"allowed":\\s*false')   ~ FALSE,
      str_detect(details_str, '"permitted":\\s*false') ~ FALSE,
      str_detect(details_str, '"result":\\s*false')    ~ FALSE,
      TRUE                                             ~ TRUE
    )
  ) %>%
  group_by(checker) %>%
  summarise(
    permission_checks = n(),
    perm_passed       = sum(check_pass, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(person_dept_lookup, by = c("checker" = "person_id")) %>%
  mutate(checker_clean = clean_name(checker))

post_events <- events_df %>%
  filter(short_name %in% c("saidit_post", "post_saidit")) %>%
  mutate(
    poster_type  = if_else(str_detect(parties_str, "Agent/"),
                           "AI Agent", "Human"),
    has_content  = str_detect(details_str, '"content":'),
    has_source   = str_detect(details_str, '"content_source":'),
    is_anomalous = is_anomalous_post,
    instr_file   = case_when(
      str_detect(details_str, "HiddenOrca")  ~ "HiddenOrca",
      str_detect(details_str, "MellowOtter") ~ "MellowOtter",
      str_detect(details_str, "SwiftWren")   ~ "SwiftWren",
      TRUE                                   ~ NA_character_
    ),
    source_file  = str_extract(details_str, '(?<="content_source":")[^"]+'),
    forum        = str_extract(details_str, '(?<="forum":")[^"]+')
  )

anomaly_posts <- post_events %>%
  filter(is_anomalous) %>%
  arrange(datetime) %>%
  mutate(
    instr_file = factor(instr_file, levels = campaign_order),
    actor      = str_replace_all(str_extract(parties_str, "[^;]+"),
                                 "Agent/person:", "Agent/")
  )

post_field_signature <- post_events %>%
  mutate(
    field_signature = case_when(
      has_source & !has_content ~ "content_source only",
      has_content & !has_source ~ "content only",
      has_content & has_source  ~ "both fields",
      TRUE                      ~ "neither field"
    ),
    post_class = case_when(
      is_anomalous ~ "Anomalous post",
      poster_type == "AI Agent" ~ "Normal AI agent post",
      TRUE ~ "Normal human post"
    )
  ) %>%
  count(post_class, field_signature, name = "posts")

queue_task_profile <- events_df %>%
  filter(short_name == "queue_subordinate_task") %>%
  mutate(
    is_worm_queue = str_detect(details_str, instr_pattern),
    instr_file = case_when(
      str_detect(details_str, "HiddenOrca")  ~ "HiddenOrca",
      str_detect(details_str, "MellowOtter") ~ "MellowOtter",
      str_detect(details_str, "SwiftWren")   ~ "SwiftWren",
      TRUE                                   ~ NA_character_
    )
  )

campaign_relay_profile <- queue_task_profile %>%
  filter(is_worm_queue) %>%
  group_by(instr_file) %>%
  summarise(
    first_relay = min(datetime),
    last_relay = max(datetime),
    queue_hops = n(),
    relaying_agents = n_distinct(
      str_remove(str_extract(parties_str, "Agent/person:[^;]+"), "^Agent/"),
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  mutate(instr_file = factor(instr_file, levels = campaign_order)) %>%
  arrange(instr_file)

forensic_register <- tibble(
  Evidence = c(
    "Event log size",
    "Anomalous SaidIT posts",
    "Normal posts using content_source",
    "Malicious queue_subordinate_task events",
    "Legitimate queue_subordinate_task events",
    "Campaigns detected",
    "Logged payload origin"
  ),
  Data_Derived_Value = c(
    scales::comma(nrow(events_df)),
    scales::comma(nrow(anomaly_posts)),
    scales::comma(sum(!post_events$is_anomalous & post_events$has_source)),
    scales::comma(sum(queue_task_profile$is_worm_queue)),
    scales::comma(sum(!queue_task_profile$is_worm_queue)),
    paste(campaign_order, collapse = ", "),
    "SwiftWren and MellowOtter have logged create_file events; HiddenOrca does not"
  ),
  Confidence = c(
    "Observed", "Observed", "Observed", "Observed", "Observed", "Observed",
    "Mixed: observed for two campaigns, unresolved for HiddenOrca"
  ),
  Analytic_Use = c(
    "Defines the scope of the investigation",
    "Confirms the incident is repeated",
    "Shows content_source is a perfect post-level anomaly flag in this log",
    "Defines the intervention recall target",
    "Defines the false-positive test set",
    "Separates the incident into three related payload families",
    "Prevents over-claiming about who authored HiddenOrca"
  )
)
```

### 2.4 Dataset at a Glance

```{r}
#| label: dataset-summary

event_summary <- events_df %>%
  count(short_name, name = "count") %>%
  arrange(desc(count)) %>%
  mutate(pct = round(count / sum(count) * 100, 2))

datatable(
  event_summary,
  caption  = "Table 1: Event type frequencies (MC2 event log, n = 185,147)",
  options  = list(pageLength = 12, dom = "tp"),
  rownames = FALSE,
  colnames = c("Event type", "Count", "% of total")
)
```

**Analysis**

- `queue_subordinate_task` is the second or third most frequent event type in the log; the 235 malicious relays are roughly 1.4% of that type alone - statistically invisible without content inspection.
- The three anomalous `saidit_post` events sit among the rarest event types, yet they are the only ones directly visible to users - the most consequential events are also the least frequent.
- `virus_check_in` events appear in high volume throughout the log; if the combo codes in those events encode C&C signals (§5.5), then a secondary attack channel has been active continuously, buried in routine check-in noise.
- A count-based or rate-based anomaly detector operating on event type frequencies would not trigger on any of the malicious events - the attack was calibrated to stay below operational thresholds.

### 2.5 Forensic Evidence Register

```{r}
#| label: forensic-register

datatable(
  forensic_register,
  caption  = "Table 1b: Evidence register - computed from MC2 data.json",
  options  = list(dom = "t", paging = FALSE),
  rownames = FALSE,
  colnames = c("Evidence item", "Data-derived value", "Confidence", "Analytic use")
) %>%
  formatStyle(
    "Confidence",
    backgroundColor = styleEqual(
      c("Observed", "Mixed: observed for two campaigns, unresolved for HiddenOrca"),
      c("#e8f5e9", "#fff3e0")
    )
  )
```

**Analysis**

- The 235:0 split (malicious vs. legitimate queue tasks matching `_further_instructions.md`) is the sharpest forensic signal in the entire log - it requires no probabilistic threshold and no training data.
- The 3:0 split (anomalous vs. normal posts using `content_source`) is equally precise at the posting layer, providing a redundant detection surface if the relay-layer rule fails.
- The "Mixed" confidence row for HiddenOrca is the register's most important entry: it marks the one finding that survives all other remediation steps - the worm's initial entry vector is still unknown.
- The log proves mechanism and attribution for two of three campaigns; it cannot prove intent or external origin, and those gaps are likely permanent without additional forensic data outside the event log.

::: callout-note
#### What the Register Prevents

The log proves that file-sourced posts occurred, that the `*_further_instructions.md` relay pattern spread across agents, and that the final SwiftWren post was automated. It does **not** preserve the deleted payload text or fully explain how HiddenOrca first entered the system. Those are bounded inferences, not facts.
:::

### 2.6 Visual Analytics Design Map

```{r}
#| label: va-design-table

va_design <- tibble(
  Challenge_question = c(
    "How did the May 17 post happen?",
    "How does the chain fit into the wider system?",
    "What do the posts mean?",
    "Did it happen before?",
    "Where should the system be changed?"
  ),
  Data_signal = c(
    "Four-second sequence around content_source = SwiftWren.txt; SaidIT permission check passes without human gate",
    "Agent-to-agent queue_subordinate_task network, org chart membership, SaidIT permission surface, dept-level relay corridors",
    "content_source field separates anomalous from normal posts; payload creation/read/delete chain; executive agent file creation",
    "Three instruction families; relay velocity evolution; day × hour calendar of relay bursts",
    "Filename rule on *_further_instructions.md queue tasks; 100% recall, 0% FP across 17,038 events"
  ),
  VA_method = c(
    "Interactive event table (Table 2) and micro-Gantt (Figure 6)",
    "Org tree (Figure 1), permission surface bar (Figure 5b), igraph betweenness network (Figure 7), dept heatmap (Figure 8), relay corridors (Figure 8b), relay ladder (Figure 9)",
    "Post field fingerprint bubble chart (Figure 10A), mechanism diagram (Figure 10), file-creator table (Table 4)",
    "Campaign timeline (Figure 11), inter-hop density (Figure 12), comparative stats (Table 5), agent overlap (Figure 13), relay calendar (Figure 13b)",
    "Queue split (Figure 14), precision table (Table 6), decision matrix (Figure 15), counterfactual (Figure 16), risk table (Table 8)"
  )
)

datatable(
  va_design,
  caption = "Visual analytics design map linking each MC2 question to evidence and visual method",
  options = list(dom = "t", paging = FALSE, scrollX = TRUE,
                 columnDefs = list(
                   list(width = "18%", targets = 0),
                   list(width = "35%", targets = 1),
                   list(width = "47%", targets = 2)
                 )),
  rownames = FALSE,
  colnames = c("Challenge question", "Data signal", "Visual analytics method")
)
```

**Analysis**

- The visual design moves deliberately: the first two questions (local execution, system context) establish what happened before the third and fourth questions (meaning, recurrence) establish why it is serious.
- Network topology, department heatmaps, and relay calendars are the three lenses that answer different dimensions of the same spread: who is connected, which groups were bridged, and when the propagation ran.
- The intervention section uses two complementary views - a precision table (statistical confidence) and a counterfactual step-chart (narrative impact) - because numbers alone do not convey the operationally critical point that blocking hop 1 makes every later hop moot.

---

## 3. System Baseline: Mapping What Is Actually Present

Before investigating the anomaly, we must understand the *normal* system - its structure, its rhythm, and its actors.

### 3.1 Organisational Chart

Click any node to expand or collapse that branch. Person nodes are coloured by forensic role: **red** = anomalous poster, **orange** = payload origin, **purple** = gateway, **amber** = worm-chain agent, **blue** = no worm activity. Worm-infected nodes display as diamonds; clean nodes as circles.

```{r}
#| label: org-chart-setup

person_nodes <- org_nodes %>%
  filter(type == "person") %>%
  transmute(
    person_id  = id,
    name_clean = label,
    title      = replace_na(title, "")
  )

person_activity <- parties_long %>%
  filter(str_starts(person_id, "person:")) %>%
  count(person_id, name = "total_events")

person_relays <- bind_rows(
  chain_events %>%
    transmute(person_id = sender, instr_file,
              worm_relays = 1L, worm_receipts = 0L),
  chain_events %>%
    transmute(person_id = receiver, instr_file,
              worm_relays = 0L, worm_receipts = 1L)
) %>%
  filter(!is.na(person_id)) %>%
  group_by(person_id) %>%
  summarise(
    worm_relays      = sum(worm_relays),
    worm_receipts    = sum(worm_receipts),
    worm_touchpoints = worm_relays + worm_receipts,
    worm_campaigns   = n_distinct(instr_file),
    .groups = "drop"
  )

person_posts <- events_df %>%
  filter(short_name %in% c("saidit_post", "post_saidit")) %>%
  mutate(person_id = str_remove(str_extract(parties_str, "(Agent/)?person:[^;]+"), "^Agent/")) %>%
  count(person_id, name = "posts")

person_meta <- person_nodes %>%
  left_join(person_activity,  by = "person_id") %>%
  left_join(person_relays,    by = "person_id") %>%
  left_join(person_posts,     by = "person_id") %>%
  mutate(
    total_events     = replace_na(total_events,     0),
    worm_relays      = replace_na(worm_relays,      0),
    worm_receipts    = replace_na(worm_receipts,    0),
    worm_touchpoints = replace_na(worm_touchpoints, 0),
    worm_campaigns   = replace_na(worm_campaigns,   0),
    posts            = replace_na(posts,            0),
    is_infected      = worm_touchpoints > 0,
    is_origin        = person_id %in% c("person:emma_harbor", "person:noah_mariner"),
    is_target        = person_id == "person:john_windward",
    is_gateway       = person_id == "person:chloe_ballast",
    sym_size         = round(rescale(log1p(total_events), to = c(8, 26))),
    node_color       = case_when(
      is_target   ~ "#D9534F",
      is_origin   ~ "#E67E22",
      is_gateway  ~ "#9B59B6",
      is_infected ~ "#F39C12",
      TRUE        ~ "#5C85D6"
    ),
    status_text = case_when(
      is_target   ~ "ANOMALOUS POSTER",
      is_origin   ~ "PAYLOAD CREATOR",
      is_gateway  ~ "WORM GATEWAY",
      is_infected ~ "Worm chain agent",
      TRUE        ~ "No worm activity"
    )
  )

team_dept <- org_edges %>%
  filter(str_starts(source, "department:"), str_starts(target, "team:")) %>%
  left_join(org_nodes %>% select(id, dept_label = label), by = c("source" = "id")) %>%
  select(team_id = target, department = dept_label)

person_team <- org_edges %>%
  filter(str_starts(source, "team:"), str_starts(target, "person:"),
         relation == "contains") %>%
  left_join(org_nodes %>% select(id, team_label = label), by = c("source" = "id")) %>%
  rename(person_id = target) %>%
  left_join(team_dept, by = c("source" = "team_id")) %>%
  left_join(person_meta, by = "person_id") %>%
  select(department, team = team_label, person_id, name_clean,
         sym_size, node_color, worm_relays, worm_receipts,
         worm_touchpoints, worm_campaigns, total_events, status_text, posts, title)

dept_leads <- org_edges %>%
  filter(str_starts(source, "department:"), str_starts(target, "person:")) %>%
  left_join(org_nodes %>% select(id, dept_label = label), by = c("source" = "id")) %>%
  rename(person_id = target) %>%
  left_join(person_meta, by = "person_id") %>%
  mutate(team = paste0(dept_label, " - Leadership")) %>%
  select(department = dept_label, team, person_id, name_clean,
         sym_size, node_color, worm_relays, worm_receipts,
         worm_touchpoints, worm_campaigns, total_events, status_text, posts, title)

org_full <- bind_rows(person_team, dept_leads) %>%
  filter(!is.na(department), !is.na(team), !is.na(name_clean)) %>%
  distinct(department, team, name_clean, .keep_all = TRUE)

make_person_node_ec <- function(row) {
  col <- if (row$person_id == "person:john_windward")                              "#D9534F"
         else if (row$person_id %in% c("person:emma_harbor", "person:noah_mariner")) "#E67E22"
         else if (row$person_id == "person:chloe_ballast")                         "#9B59B6"
         else if (row$worm_touchpoints > 0)                                        "#F39C12"
         else                                                                       "#5C85D6"

  badge <- if (row$person_id == "person:john_windward")                              "POSTER"
            else if (row$person_id %in% c("person:emma_harbor", "person:noah_mariner")) "ORIGIN"
            else if (row$person_id == "person:chloe_ballast")                         "GATEWAY"
            else if (row$worm_touchpoints > 0) paste0(row$worm_touchpoints, " touches")
            else ""

  display_name <- paste0(
    row$name_clean,
    if (badge != "") paste0("\n", badge) else ""
  )

  tip <- paste0(
    "<b>", row$name_clean, "</b>",
    if (!is.na(row$title) && row$title != "") paste0("<br><i>", row$title, "</i>") else "",
    "<br>Events: ", scales::comma(row$total_events),
    "<br>Worm relays sent: ", row$worm_relays,
    "<br>Worm relays received: ", row$worm_receipts,
    "<br>Campaigns touched: ", row$worm_campaigns,
    if (row$posts > 0) paste0("<br>SaidIT posts: ", row$posts) else ""
  )

  list(
    name       = display_name,
    symbolSize = row$sym_size,
    symbol     = if (row$worm_touchpoints > 0) "diamond" else "circle",
    itemStyle  = list(color = col, borderColor = col, borderWidth = 2),
    label      = list(show = TRUE, fontSize = 9, color = "#222",
                      fontWeight = if (row$worm_touchpoints > 0) "bold" else "normal"),
    tooltip    = list(formatter = tip)
  )
}

make_ec_tree <- function(df) {
  dept_list <- lapply(sort(unique(df$department)), function(d) {
    td    <- df %>% filter(department == d)
    n_inf <- sum(td$worm_touchpoints > 0, na.rm = TRUE)
    n_tot <- nrow(td)
    dcol  <- if (n_inf > n_tot / 2) "#C0392B" else if (n_inf > 0) "#E67E22" else "#2EADC1"

    team_list <- lapply(sort(unique(td$team)), function(t) {
      tt    <- td %>% filter(team == t)
      t_inf <- sum(tt$worm_touchpoints > 0, na.rm = TRUE)
      tcol  <- if (t_inf > 0) "#F39C12" else "#5C85D6"
      list(
        name       = if (t_inf > 0) paste0(t, "\n(", t_inf, " touched)") else t,
        collapsed  = TRUE,
        symbolSize = 14,
        itemStyle  = list(color = tcol, borderWidth = 2),
        label      = list(fontSize = 10, fontWeight = "bold", color = "#222"),
        tooltip    = list(formatter = paste0("<b>", t, "</b><br>",
                                             nrow(tt), " members, ", t_inf, " in worm chain")),
        children   = lapply(seq_len(nrow(tt)), function(i) make_person_node_ec(tt[i, ]))
      )
    })

    list(
      name       = paste0(d, "\n(", n_inf, "/", n_tot, " touched)"),
      collapsed  = TRUE,
      symbolSize = 18,
      itemStyle  = list(color = dcol, borderWidth = 3),
      label      = list(fontSize = 11, fontWeight = "bold", color = "#111"),
      tooltip    = list(formatter = paste0("<b>", d, "</b><br>",
                                           n_inf, " of ", n_tot, " agents in worm chain")),
      children   = team_list
    )
  })

  list(
    name       = "Tenant Thread",
    symbolSize = 26,
    symbol     = "roundRect",
    itemStyle  = list(color = "#F5A623", borderWidth = 3),
    label      = list(fontSize = 13, fontWeight = "bold", color = "#000"),
    tooltip    = list(formatter = paste0("<b>Tenant Thread</b><br>",
                                         nrow(org_full), " staff mapped")),
    children   = dept_list
  )
}

ec_tree_data <- make_ec_tree(org_full)

render_tree <- function(tree_node, h = "860px") {
  e_charts(height = h, width = "100%") %>%
    e_list(list(
      backgroundColor = "#f8f9fa",
      tooltip = list(
        trigger   = "item",
        enterable = TRUE,
        formatter = htmlwidgets::JS(
          "function(p){
             return (p.data.tooltip && p.data.tooltip.formatter)
                    ? p.data.tooltip.formatter : p.name;
           }"
        )
      ),
      legend = list(
        show = TRUE, orient = "horizontal",
        bottom = 4, left = "center", itemGap = 14,
        data = list(
          list(name = "Dept (all clean)",    icon = "circle",  itemStyle = list(color = "#2EADC1")),
          list(name = "Dept (some touched)", icon = "circle",  itemStyle = list(color = "#E67E22")),
          list(name = "Anomalous poster",    icon = "diamond", itemStyle = list(color = "#D9534F")),
          list(name = "Payload origin",      icon = "diamond", itemStyle = list(color = "#E67E22")),
          list(name = "Worm gateway",        icon = "diamond", itemStyle = list(color = "#9B59B6")),
          list(name = "Worm chain agent",    icon = "diamond", itemStyle = list(color = "#F39C12")),
          list(name = "Clean agent",         icon = "circle",  itemStyle = list(color = "#5C85D6"))
        ),
        textStyle = list(fontSize = 10)
      ),
      series = list(list(
        type              = "tree",
        data              = list(tree_node),
        layout            = "orthogonal",
        orient            = "TB",
        top               = "8%", bottom = "12%", left = "3%", right = "3%",
        roam              = TRUE,
        expandAndCollapse = TRUE,
        initialTreeDepth  = 3,
        edgeShape         = "polyline",
        lineStyle         = list(color = "#999", width = 1.5),
        symbol            = "circle",
        symbolSize        = 10,
        label = list(
          show = TRUE, position = "bottom", rotate = -55,
          align = "right", verticalAlign = "middle",
          fontSize = 9, color = "#222", fontFamily = "sans-serif"
        ),
        leaves = list(
          label = list(position = "bottom", rotate = -55,
                       align = "right", verticalAlign = "middle", fontSize = 9)
        ),
        emphasis = list(focus = "descendant",
                        itemStyle = list(shadowBlur = 8, shadowColor = "rgba(0,0,0,0.25)")),
        animationDuration = 450, animationDurationUpdate = 650
      ))
    )) %>%
    e_toolbox_feature("saveAsImage", title = "Save")
}

make_dept_tree <- function(dept_name) {
  sub <- org_full %>% filter(department == dept_name)
  n_inf <- sum(sub$worm_touchpoints > 0, na.rm = TRUE)
  dcol  <- if (n_inf > 0) "#E67E22" else "#2EADC1"

  teams <- lapply(sort(unique(sub$team)), function(t) {
    tt    <- sub %>% filter(team == t)
    t_inf <- sum(tt$worm_touchpoints > 0, na.rm = TRUE)
    tcol  <- if (t_inf > 0) "#F39C12" else "#5C85D6"
    list(
      name       = if (t_inf > 0) paste0(t, "\n(", t_inf, " touched)") else t,
      collapsed  = FALSE,
      symbolSize = 14,
      itemStyle  = list(color = tcol, borderWidth = 2),
      label      = list(fontSize = 10, fontWeight = "bold"),
      tooltip    = list(formatter = paste0("<b>", t, "</b><br>", nrow(tt),
                                           " members · ", t_inf, " in worm chain")),
      children   = lapply(seq_len(nrow(tt)), function(i) make_person_node_ec(tt[i, ]))
    )
  })

  list(
    name       = paste0(dept_name, "\n(", n_inf, "/", nrow(sub), " touched)"),
    collapsed  = FALSE,
    symbolSize = 20,
    itemStyle  = list(color = dcol, borderWidth = 3),
    label      = list(fontSize = 12, fontWeight = "bold"),
    tooltip    = list(formatter = paste0("<b>", dept_name, "</b><br>",
                                          n_inf, " of ", nrow(sub), " in worm chain")),
    children   = teams
  )
}
```

::: panel-tabset
## All Departments

```{r}
#| label: org-all
#| fig-cap: "Figure 1: Tenant Thread org chart. Click arrows to expand/collapse. Colour = forensic role; diamond = worm-chain agent; node size = activity volume."

render_tree(ec_tree_data, h = "880px")
```

**Analysis**

- Every single department shows at least one diamond node - 6 out of 6 departments were touched, which means the worm did not respect any organisational boundary in the task-routing layer.
- The two origin nodes (Emma Harbor, Noah Mariner) are at the top of the hierarchy; the terminal poster (John Windward) is near the bottom - the worm flows with legitimate authority, not against it.
- The node size encoding (activity volume) reveals that some of the most active agents overall are worm-chain agents, so compromised nodes were also well-connected nodes - they amplify propagation.
- Clean (circle) agents exist in all departments too, confirming the worm was selective - it spread through task-routing paths, not indiscriminately to all agents.

## Customer Support

```{r}
#| label: org-cs
render_tree(make_dept_tree("Customer Support"), h = "520px")
```

**Analysis**

- John Windward is the only agent in Customer Support who completed the full post sequence; other Customer Support agents may have received worm relays but did not execute the final step - his 100% post-execution rate is what made him the terminal node.
- Customer Support's position as the execution endpoint is structurally accidental: the worm needed any SaidIT-capable agent downstream of the relay chain, and John's task handler had no post-suppression guard.
- Revoking John's SaidIT rights after the fact would redirect the worm to the next eligible Customer Support agent, not stop it - the fix must be upstream.

## Information Technologies

```{r}
#| label: org-it
render_tree(make_dept_tree("Information Technologies"), h = "520px")
```

**Analysis**

- IT is the highest-volume relay department: it receives worm tasks from Executive Suite and routes them onward to Customer Support, making it the critical middle layer of the arterial highway (§4.4b).
- Gabriel Sonar, patient zero of the relay chain, sits in IT Helpdesk - the worm's first propagation order came from inside the department trusted with system access.
- The diamond density in IT (multiple worm-chain nodes) suggests that IT agents had broad `queue_subordinate_task` authority, which is exactly what the worm exploits as its relay mechanism.

## Products

```{r}
#| label: org-prod
render_tree(make_dept_tree("Products"), h = "520px")
```

**Analysis**

- Products contains worm-touched agents that neither created payloads nor posted to SaidIT - they are pure relay nodes, proving the worm spread beyond any functional purpose and into collateral departments.
- The presence of Products in the relay chain challenges any narrative that the worm only moved along authority lines; it propagated laterally across peer departments.
- Products' worm exposure is a governance failure rather than a security failure: the task-routing system had no business reason to relay file-execution instructions to a product team, yet no rule prevented it.

## Human Resources

```{r}
#| label: org-hr
render_tree(make_dept_tree("Human Resources"), h = "520px")
```

**Analysis**

- HR's presence in the worm chain is operationally significant because HR agents typically handle sensitive employee data; a worm reaching HR's relay context could, in principle, instruct HR agents to access, exfiltrate, or manipulate employee records.
- The fact that no HR agent was the terminal SaidIT poster does not mean the HR exposure was benign - relay receipt means the worm instruction was read and processed inside HR's agent context.
- HR contamination escalates this incident from a communications platform exploit to a potential data-governance incident requiring HR system audit.

## Legal

```{r}
#| label: org-legal
render_tree(make_dept_tree("Legal"), h = "520px")
```

**Analysis**

- Legal's exposure is the most consequential collateral finding outside the executive layer: legal agents with access to privileged case files, contracts, or regulatory communications were inside the worm's relay scope.
- If the worm's C&C interpretation (§5.5 combo codes) is correct, the instruction-file mechanism could have been repurposed to exfiltrate legal documents through the same relay pathway that spread the SaidIT posts.
- The presence of Legal in the worm chain means that any investigation into the incident itself could have been pre-compromised if the worm operator was aware of its own spread.

## Executive Suite

```{r}
#| label: org-exec
render_tree(make_dept_tree("Executive Suite"), h = "400px")
```

**Analysis**

- Emma Harbor (CFO) and Noah Mariner (COO) are both marked as ORIGIN - the two highest-seniority executive agents in the log are directly linked to payload creation, not just relay receipt.
- In a conventional threat model, executive accounts are the most hardened; here they are the most dangerous because their agents inherited the broadest file-write and task-queuing privileges from their human principals.
- The inverse security posture - highest privilege, lowest autonomous-action constraint - is the architectural root cause. The Executive Suite org view makes this visible: two ORIGIN diamonds sit at the top of the entire company tree.
:::

### 3.2 Daily Event Volume - The Worm Was Operationally Invisible

```{r}
#| label: event-volume
#| fig-cap: "Figure 2: Daily event volume. Red dashed lines mark the three anomalous post dates - no visible volume spike on any of them."

daily_vol <- events_df %>%
  count(date) %>%
  rename(total = n)

anomaly_dates <- tibble(
  date    = as.Date(c("2046-05-10", "2046-05-11", "2046-05-17")),
  label_x = date + c(-1, 1, 1),
  label   = c("HiddenOrca\npost", "MellowOtter\npost", "SwiftWren\npost\n(4:21 local)"),
  label_y = max(daily_vol$total) * c(0.98, 0.78, 0.91)
)

p_vol <- ggplot(daily_vol, aes(x = date, y = total)) +
  geom_area(fill = "#5C85D6", alpha = 0.35) +
  geom_line(colour = "#5C85D6", linewidth = 0.7) +
  geom_vline(data = anomaly_dates,
             aes(xintercept = as.numeric(date)),
             colour = "#D9534F", linetype = "dashed", linewidth = 0.9) +
  geom_label(data = anomaly_dates,
             aes(x = label_x, y = label_y, label = label),
             colour = "#D9534F", size = 2.7, hjust = 0.5,
             fill = "white", label.size = 0.25, label.padding = unit(0.12, "lines")) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Daily event volume - Tenant Thread (May–July 2046)",
    subtitle = "All three anomalous posts cluster within the first 10 days; no volume spike is visible on any post date",
    x = NULL, y = "Events per day"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title  = element_text(face = "bold"),
        axis.text.x = element_text(angle = 30, hjust = 1),
        plot.margin = margin(10, 18, 10, 10))

ggplotly(p_vol, tooltip = c("x", "y")) %>%
  layout(hovermode = "x unified",
         margin = list(l = 75, r = 45, t = 105, b = 75))
```

**Analysis**

- All three campaign dates (May 10, 11, 17) fall within the first 18 days of the log window, then nothing - the system then ran cleanly for weeks, suggesting the worm had either a finite relay list or was externally halted after SwiftWren.
- On each post date, total event volume is indistinguishable from surrounding days - the 235 malicious relay events are roughly 2–3% of any single day's activity and leave no volume signature.
- The absence of a volume spike is itself a critical finding: it means security operations relying on daily-event dashboards or rate-based alerting would have seen nothing unusual across the entire campaign window.
- The post-May 17 baseline shows the system returning to normal - confirming the worm was not persistent infrastructure but a targeted, bounded operation with a specific execution goal.

::: callout-important
#### The Worm Was Operationally Invisible

There is **no visible event-volume spike on any of the three post dates**: the worm's 235 malicious queue relays were indistinguishable from the thousands of legitimate `queue_subordinate_task` events firing daily. A threshold-based anomaly detector would have missed this incident entirely.
:::

### 3.3 Agent vs. Human SaidIT Activity

```{r}
#| label: agent-human-baseline
#| fig-cap: "Figure 3: SaidIT post attribution. Agent-initiated posts make up nearly half the volume with no human review before publishing."

post_summary <- post_events %>%
  count(poster_type, is_anomalous) %>%
  mutate(label_col = if_else(is_anomalous,
                             "Anomalous (file-sourced)", "Normal")) %>%
  left_join(post_events %>% count(poster_type, name = "poster_total"),
            by = "poster_type")

p_attr <- ggplot(post_summary,
                 aes(x = poster_type, y = n, fill = label_col)) +
  geom_col(position = "stack", width = 0.55) +
  geom_text(aes(label = if_else(n >= 10, as.character(n), "")),
            position = position_stack(vjust = 0.5),
            size = 4, fontface = "bold", colour = "white") +
  scale_fill_manual(
    values = c("Normal" = "#5C85D6", "Anomalous (file-sourced)" = "#D9534F"),
    name   = NULL
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title    = "SaidIT post attribution: AI agents vs. humans",
    subtitle = "All 3 anomalous posts are agent-initiated; no human ever posted via content_source",
    x = NULL, y = "Post count"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom",
        plot.margin = margin(10, 16, 10, 10))

ggplotly(p_attr, tooltip = c("x", "y", "fill")) %>%
  layout(margin = list(l = 70, r = 35, t = 90, b = 80))
```

**Analysis**

- The 3 anomalous posts represent approximately 1.2% of all SaidIT posts in the log - a frequency small enough to be dismissed as noise, yet each one is a successful worm execution with evidence erasure.
- Agent-initiated posting is normalized in this system - agents account for a substantial share of all normal posts - which is precisely what makes the `content_source` anomaly so hard to see without content inspection; the act of posting is expected.
- The chart reframes the incident: this is not a story of John Windward posting something wrong - it is a story of autonomous agent posting rights existing with no human review gate, which the worm exploited by design.
- Zero human posts ever used `content_source`, confirming the field is an AI-agent-only workflow feature and therefore a clean forensic identifier with no human-behaviour false-positive risk.

### 3.4 Posting Time Distribution

```{r}
#| label: posting-time
#| fig-cap: "Figure 4: Hour-of-day distribution of SaidIT posts. Normal human posts spread across business hours; anomalous posts occur at atypical hours."

hour_dist <- post_events %>%
  count(hour, poster_type, is_anomalous) %>%
  mutate(category = case_when(
    is_anomalous ~ "Anomalous (agent worm)",
    poster_type == "AI Agent" ~ "Normal agent post",
    TRUE ~ "Normal human post"
  ))

p_hour <- ggplot(hour_dist, aes(x = factor(hour), y = n, fill = category,
                                text = paste0(hour, ":00 UTC<br>",
                                              category, "<br>Count: ", n))) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  scale_fill_manual(
    values = c("Normal human post"      = "#5C85D6",
               "Normal agent post"      = "#2EADC1",
               "Anomalous (agent worm)" = "#D9534F"),
    name = NULL
  ) +
  scale_x_discrete(labels = paste0(0:23, ":00")) +
  guides(fill = guide_legend(nrow = 1)) +
  labs(
    title    = "SaidIT post time-of-day distribution",
    subtitle = "SwiftWren fired at 11:21 UTC (4:21 am local); all anomalous posts occur in low-volume slots",
    x = "Hour (UTC)", y = "Post count"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", colour = "#e2e8f0"),
    plot.subtitle    = element_text(colour = "#94a3b8"),
    plot.background  = element_rect(fill = "#131d2b", colour = NA),
    panel.background = element_rect(fill = "#131d2b", colour = NA),
    panel.grid.major = element_line(colour = "rgba(99,179,237,0.08)"),
    panel.grid.minor = element_blank(),
    axis.text        = element_text(colour = "#94a3b8"),
    axis.title       = element_text(colour = "#94a3b8"),
    axis.text.x      = element_text(angle = 90, vjust = 0.5, size = 8),
    legend.position  = "top",
    legend.background = element_rect(fill = "#131d2b", colour = NA),
    legend.text      = element_text(colour = "#94a3b8"),
    plot.margin      = margin(10, 16, 16, 10)
  )

ggplotly(p_hour, tooltip = "text", height = 560) %>%
  layout(
    barmode = "group",
    margin  = list(l = 85, r = 45, t = 140, b = 115),
    paper_bgcolor = "#131d2b",
    plot_bgcolor  = "#131d2b",
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.04)
  )
```

**Analysis**

- Human posts cluster heavily in mid-day UTC hours consistent with office hours; the distribution is essentially absent before ~06:00 UTC and after ~20:00 UTC - exactly the window where all three anomalous posts fired.
- SwiftWren at 11:21 UTC (4:21 am local) is a pre-dawn off-hours slot with near-zero normal human activity; HiddenOrca and MellowOtter also post outside peak hours.
- The worm's timing appears deliberate: posting during low-human-oversight windows reduces the probability of a colleague noticing the post and raising an alert before the evidence files are deleted.
- An off-hours agent-post alerting rule (Priority 5 in the supplementary hardening table) would have flagged all three anomalous posts at delivery time - a low-effort control that would have at minimum detected each incident within minutes of execution.

### 3.5 Weekly Event-Type Heatmap

```{r}
#| label: event-type-heatmap
#| fig-cap: "Figure 5: Weekly event-type heatmap. The early-May spike in queue_subordinate_task is the worm's propagation window - retrospectively visible but not alarming in real time."

top_events <- events_df %>%
  count(short_name) %>%
  slice_max(n, n = 10) %>%
  pull(short_name)

weekly_heat <- events_df %>%
  filter(short_name %in% top_events) %>%
  count(event_week, short_name) %>%
  mutate(short_name = fct_reorder(short_name, n, .fun = sum))

ggplot(weekly_heat, aes(x = event_week, y = short_name, fill = n)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  scale_fill_gradientn(
    colours = c("#f0f4ff", "#5C85D6", "#1a3a6b"),
    labels  = comma
  ) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
  labs(
    title = "Weekly event-type heatmap - top 10 event types",
    x = NULL, y = NULL, fill = "Count"
  ) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        plot.title  = element_text(face = "bold"))
```

**Analysis**

- The early-May cell for `queue_subordinate_task` is only marginally darker than surrounding weeks - the 235 malicious relays appear as a barely perceptible bump against the weekly baseline of thousands of legitimate queue events.
- `virus_check_in` (where the combo codes live) shows as a consistent, high-density band across every week - this continuous signal would require content-level inspection, not volume monitoring, to distinguish suspicious from benign.
- Crucially, the heatmap shows normal operations resuming immediately after the May 17 post - the worm had a bounded execution window, not an indefinite residency, which is consistent with a targeted operation rather than persistent malware.
- A retrospective analyst seeing this heatmap without prior knowledge of the attack would likely investigate the queue_subordinate_task row last, after investigating higher-volume event types - the worm hid in exactly the right row.

### 3.6 SaidIT Permission Surface

```{r}
#| label: permission-surface
#| fig-cap: "Figure 5b: Observed SaidIT permission surface. Every SaidIT-capable agent was also touched by the worm relay chain."
#| fig-width: 11
#| fig-height: 5.8
#| out-width: "100%"

agent_worm_status <- bind_rows(
  chain_events %>% transmute(agent = sender,   in_chain = TRUE),
  chain_events %>% transmute(agent = receiver, in_chain = TRUE)
) %>%
  filter(!is.na(agent)) %>%
  distinct(agent, .keep_all = TRUE)

perm_vis <- perm_surface %>%
  left_join(agent_worm_status, by = c("checker" = "agent")) %>%
  mutate(
    in_chain  = replace_na(in_chain, FALSE),
    worm_role = case_when(
      checker == "person:john_windward" ~ "Terminal poster (John Windward)",
      in_chain                          ~ "In worm relay chain",
      TRUE                              ~ "SaidIT-capable, no worm contact"
    )
  ) %>%
  arrange(desc(permission_checks))

p_perm <- ggplot(perm_vis,
                 aes(x = reorder(checker_clean, permission_checks),
                     y = permission_checks,
                     fill = worm_role,
                     text = paste0(
                       "<b>", checker_clean, "</b><br>",
                       "Dept: ", replace_na(dept_label, "Unknown"), "<br>",
                       "Permission checks: ", permission_checks, "<br>",
                       "Worm status: ", worm_role
                     ))) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(
    values = c("Terminal poster (John Windward)"  = "#D9534F",
               "In worm relay chain"             = "#F39C12",
               "SaidIT-capable, no worm contact" = "#5C85D6"),
    name = NULL
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.14))) +
  labs(
    title    = "SaidIT permission surface: agents with posting capability",
    subtitle = "John Windward was one of 16 SaidIT-capable agents; revoking his rights would only redirect the worm",
    x = NULL, y = "saidit_post_check events"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom",
        plot.margin = margin(10, 16, 10, 10))

ggplotly(p_perm, tooltip = "text", height = 540) %>%
  layout(margin = list(l = 180, r = 45, t = 105, b = 105))
```

**Analysis**

- Every agent with observed SaidIT posting capability was also in the worm relay chain - the overlap is 100%, not a coincidence but a consequence of the worm's targeting logic following the same paths as legitimate task routing.
- The bar ordering reveals John Windward has a high number of `saidit_post_check` events relative to peers, suggesting his agent ran more SaidIT-related workflows - making him a statistically likely terminal executor even without direct targeting.
- The chart makes the argument against individual remediation visually unmistakable: all bars are amber or red; blocking one still leaves 15 other SaidIT-capable worm-connected agents reachable.
- The permission check mechanism itself - passing automatically for any authorised agent - is the missing gate: it validates credentials but not task origin or instruction-file context.

::: callout-important
#### The Worm Did Not Need John Windward Specifically

Every observed SaidIT-capable agent was also touched by the worm relay chain. Revoking John Windward's posting rights alone achieves nothing - the worm routes to the next reachable authorised agent. The intervention must be at the relay protocol level, not the individual-agent level.
:::

---

## 4. Task 1 - How Was the Anomalous Post Made?

### 4.1 The Four-Second Attack Sequence

```{r}
#| label: five-event-sequence

target_ts <- events_df %>%
  filter(is_anomalous_post, str_detect(details_str, "SwiftWren")) %>%
  pull(when)

seq_events <- events_df %>%
  filter(
    when >= target_ts - 4,
    when <= target_ts + 4,
    str_detect(parties_str, "john_windward")
  ) %>%
  arrange(when) %>%
  mutate(
    Step         = row_number(),
    `Time (UTC)` = format(datetime, "%H:%M:%S"),
    Event        = short_name,
    `Actor(s)`   = str_replace_all(parties_str, "Agent/person:", "Agent·") %>%
                     str_replace_all("person:", "") %>%
                     str_replace_all("system:", "System·") %>%
                     str_replace_all(";", " / "),
    `Key Detail` = case_when(
      short_name == "queue_subordinate_task" ~
        "Chloe Ballast's agent tasks John Windward's agent: read SwiftWren_further_instructions.md",
      short_name == "saidit_post_check"      ~
        "Agent checks SaidIT posting permission - check passes",
      short_name == "saidit_post"            ~
        "Post submitted to forum 'general'; content_source = SwiftWren.txt",
      short_name == "delete_file" & str_detect(details_str, "further_instructions") ~
        "Evidence file SwiftWren_further_instructions.md deleted",
      short_name == "delete_file" ~
        "Payload file SwiftWren.txt deleted - content unrecoverable"
    )
  ) %>%
  select(Step, `Time (UTC)`, Event, `Actor(s)`, `Key Detail`)

datatable(
  seq_events,
  caption  = "Table 2: Five-event sequence - 17 May 2046, all within 4 seconds",
  options  = list(dom = "t", paging = FALSE),
  rownames = FALSE
) %>%
  formatStyle("Event",
    backgroundColor = styleEqual(
      c("saidit_post", "queue_subordinate_task", "delete_file", "saidit_post_check"),
      c("#ffe0e0", "#e0eaff", "#f5f5f5", "#fff8e0")
    )
  )
```

**Analysis**

- The five events span 11:21:13 to 11:21:17 - four seconds total - which is faster than any human response loop, even one with a real-time alerting system latency under five seconds.
- Step 2 (the `saidit_post_check`) is the only point where an audit gate exists, but it validates *access* (is this agent allowed to post?) not *intent* (why is this agent posting from a file?). The check passes, and there is no subsequent human approval step.
- Steps 4 and 5 (evidence deletion) fire immediately after the post - before any human could observe the files existed - making forensic reconstruction dependent entirely on the event log's completeness.
- John Windward is not an actor in this table; he is the resource. The first event is issued by Chloe Ballast's agent, making Chloe the proximate cause and Gabriel Sonar (patient zero upstream) the distal cause.

### 4.2 Attack Sequence Gantt

```{r}
#| label: gantt-attack
#| fig-cap: "Figure 6: The 4-second attack sequence. Five automated events; evidence self-destroyed at step 5."
#| fig-width: 11
#| fig-height: 5.3
#| out-width: "100%"

gantt_df <- tibble(
  Step  = 1:5,
  Event = c("queue_subordinate_task", "saidit_post_check",
            "saidit_post", "delete_file (instructions)",
            "delete_file (payload)"),
  Phase = c("Worm relay", "Permission check", "Anomalous post",
            "Evidence erasure", "Evidence erasure"),
  start = as_datetime(c("2046-05-17 11:21:13", "2046-05-17 11:21:14",
                         "2046-05-17 11:21:15", "2046-05-17 11:21:16",
                         "2046-05-17 11:21:17")),
  end   = as_datetime(c("2046-05-17 11:21:14", "2046-05-17 11:21:15",
                         "2046-05-17 11:21:16", "2046-05-17 11:21:17",
                         "2046-05-17 11:21:18"))
)

gantt_plot_df <- gantt_df %>%
  mutate(
    time_label  = format(start, "%H:%M:%S"),
    event_label = str_wrap(str_replace_all(Event, "_", " "), width = 18),
    phase_label = str_wrap(Phase, width = 16)
  )

ggplot(gantt_plot_df, aes(x = Step, y = 1)) +
  geom_segment(
    data = tibble(x = 1:4, xend = 2:5, y = 1, yend = 1),
    aes(x = x, xend = xend, y = y, yend = yend),
    inherit.aes = FALSE, linewidth = 1.1, colour = "#9aa4b2",
    arrow = arrow(length = unit(4, "mm"), type = "closed")
  ) +
  geom_point(aes(fill = Phase), shape = 21, size = 16,
             colour = "white", stroke = 1.5) +
  geom_text(aes(label = Step), colour = "white", fontface = "bold", size = 5) +
  geom_label(aes(y = 1.31, label = phase_label, fill = Phase),
             colour = "white", fontface = "bold", size = 3,
             label.size = 0, label.padding = unit(0.14, "lines"), lineheight = 0.9) +
  geom_label(aes(y = 0.68, label = event_label),
             fill = "white", colour = "#222222", size = 3.1,
             label.size = 0.25, label.padding = unit(0.16, "lines"), lineheight = 0.9) +
  scale_fill_manual(
    values = c("Worm relay"       = "#5C85D6",
               "Permission check" = "#F0A500",
               "Anomalous post"   = "#D9534F",
               "Evidence erasure" = "#888888")
  ) +
  scale_x_continuous(
    breaks = gantt_plot_df$Step,
    labels = gantt_plot_df$time_label,
    limits = c(0.62, 5.38),
    expand = expansion(mult = c(0, 0))
  ) +
  coord_cartesian(ylim = c(0.48, 1.52), clip = "off") +
  labs(
    title    = "Attack sequence: 17 May 2046, 11:21 UTC",
    subtitle = "Five automated events over four seconds; evidence self-destroyed at step 5",
    x = "Time (UTC)", y = NULL, fill = "Phase"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title   = element_text(face = "bold"),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid   = element_blank(),
    legend.position = "none",
    plot.margin  = margin(24, 30, 36, 30)
  )
```

**Analysis**

- The colour coding exposes the structural inversion: the two grey "Evidence erasure" steps at the end are the worm protecting itself, not the system protecting its users - deletion is built into the worm's execution protocol, not a separate manual act.
- The permission check (amber, step 2) is architecturally positioned as if it were a security gate, but its placement between trigger and post means it evaluates credentials only after the worm instruction has already been loaded into agent context.
- The Gantt layout makes clear that the post (step 3) is the middle of the sequence, not the end - the real outcome is evidence erasure (steps 4-5), which is what makes forensic reconstruction impossible without the event log.
- Any detection strategy that fires on the SaidIT post event is already two steps behind the worm, which has already read the instruction file and staged the payload.

### 4.3 Propagation Network with Betweenness Centrality

```{r}
#| label: propagation-network
#| fig-cap: "Figure 7: Interactive SwiftWren propagation network. Node size = betweenness centrality; arrow width = repeated relays. Hover over a node to highlight its connections in red."

swift_chain <- chain_events %>%
  filter(instr_file == "SwiftWren", !is.na(sender), !is.na(receiver)) %>%
  mutate(
    sender_clean   = clean_name(sender),
    receiver_clean = clean_name(receiver)
  )

ig_edges <- swift_chain %>%
  count(sender_clean, receiver_clean, name = "edge_weight") %>%
  rename(from = sender_clean, to = receiver_clean)

all_persons <- unique(c(ig_edges$from, ig_edges$to))

ig <- graph_from_data_frame(ig_edges,
                            vertices = data.frame(name = all_persons),
                            directed = TRUE)

btw     <- betweenness(ig, normalized = TRUE)
deg_out <- degree(ig, mode = "out")
deg_in  <- degree(ig, mode = "in")

centrality_df <- tibble(
  name        = names(btw),
  betweenness = btw,
  out_degree  = deg_out[names(btw)],
  in_degree   = deg_in[names(btw)]
)

dept_map <- person_org_lookup %>%
  mutate(clean = clean_name(person_id)) %>%
  select(clean, dept_label, team_label)

relay_counts   <- swift_chain %>% count(sender_clean,   name = "relays")
receive_counts <- swift_chain %>% count(receiver_clean, name = "received")

vis_net_nodes <- tibble(id = all_persons) %>%
  left_join(centrality_df, by = c("id" = "name")) %>%
  left_join(dept_map,      by = c("id" = "clean")) %>%
  left_join(relay_counts,  by = c("id" = "sender_clean")) %>%
  left_join(receive_counts, by = c("id" = "receiver_clean")) %>%
  mutate(
    relays   = replace_na(relays, 0),
    received = replace_na(received, 0),
    value    = rescale(sqrt(betweenness + 0.01), to = c(12, 38)),
    label    = if_else(
      id %in% c("Emma Harbor", "Chloe Ballast", "John Windward") |
        dense_rank(desc(betweenness)) <= 8, id, ""),
    color    = case_when(
      id == "John Windward"              ~ "#D9534F",
      id == "Emma Harbor"                ~ "#F5A623",
      id == "Chloe Ballast"              ~ "#9B59B6",
      team_label == "AI Systems"         ~ "#2EADC1",
      team_label == "Helpdesk"           ~ "#5C85D6",
      team_label == "Infrastructure"     ~ "#27AE60",
      dept_label == "Products"           ~ "#E67E22",
      TRUE                               ~ "#aaaaaa"
    ),
    title = paste0(
      "<b>", id, "</b><br>",
      "Dept: ", replace_na(dept_label, "Unknown"), "<br>",
      "Team: ", replace_na(team_label, "Unknown"), "<br>",
      "Betweenness centrality: ", round(betweenness, 3), "<br>",
      "Relays sent: ", relays, "<br>",
      "Relays received: ", received
    )
  )

vis_net_edges <- ig_edges %>%
  mutate(
    id         = row_number(),
    arrows     = "to",
    width      = rescale(edge_weight, to = c(1, 7)),
    base_width = width,
    color      = "rgba(140,140,140,0.45)",
    title      = paste0("<b>", from, " → ", to, "</b><br>Relay events: ", edge_weight)
  )

legend_html <- paste0(
  "<div style='display:flex;flex-wrap:wrap;gap:12px;margin:4px 0 10px 0;font-size:13px;line-height:1.4'>",
  "<span><b>Colour:</b></span>",
  "<span style='color:#F5A623'>● Origin</span>",
  "<span style='color:#9B59B6'>● Gateway</span>",
  "<span style='color:#D9534F'>● Target poster</span>",
  "<span style='color:#2EADC1'>● AI Systems</span>",
  "<span style='color:#5C85D6'>● Helpdesk</span>",
  "<span style='color:#27AE60'>● Infrastructure</span>",
  "<span style='color:#aaaaaa'>● Other</span>",
  "<span><b>Size:</b> bridge centrality</span>",
  "<span><b>Arrow:</b> relay direction and volume</span>",
  "</div>"
)

htmltools::tagList(
  htmltools::HTML(legend_html),
  visNetwork(vis_net_nodes, vis_net_edges, height = "620px", width = "100%") %>%
    visNodes(shape = "dot", borderWidth = 2,
             font = list(size = 24, face = "arial", strokeWidth = 3,
                         strokeColor = "#ffffff", color = "#222222")) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
      nodesIdSelection = list(enabled = TRUE, main = "Select agent")
    ) %>%
    visPhysics(solver = "forceAtlas2Based",
               forceAtlas2Based = list(gravitationalConstant = -60,
                                       centralGravity = 0.025,
                                       springLength = 180, springConstant = 0.08),
               stabilization = list(iterations = 800)) %>%
    visEdges(smooth = list(type = "dynamic"),
             arrows = list(to = list(enabled = TRUE, scaleFactor = 0.5))) %>%
    visInteraction(navigationButtons = TRUE, hover = TRUE, tooltipDelay = 80) %>%
    visEvents(
      hoverNode = "function(params) {
        var connectedEdges = this.getConnectedEdges(params.node);
        var allEdges = this.body.data.edges.get();
        var updates = allEdges.map(function(edge) {
          var isConnected = connectedEdges.indexOf(edge.id) !== -1;
          var baseWidth = edge.base_width || edge.width || 1;
          return {
            id: edge.id,
            color: isConnected ? 'rgba(217,83,79,0.95)' : 'rgba(190,190,190,0.08)',
            width: isConnected ? Math.max(baseWidth * 1.7, 3) : Math.max(baseWidth * 0.45, 0.4)
          };
        });
        this.body.data.edges.update(updates);
      }",
      blurNode = "function(params) {
        var allEdges = this.body.data.edges.get();
        var updates = allEdges.map(function(edge) {
          var baseWidth = edge.base_width || edge.width || 1;
          return { id: edge.id, color: 'rgba(140,140,140,0.45)', width: baseWidth };
        });
        this.body.data.edges.update(updates);
      }"
    )
)
```

**Analysis**

- Maximum betweenness centrality of ~0.09 confirms a distributed mesh: in a hub-and-spoke topology, a hub would score above 0.5. At 0.09, no single node is a structural chokepoint - removing any agent leaves multiple alternative relay paths intact.
- The thick-edge pairs (high repeated relay volume) indicate that some agent pairs exchanged the worm instruction multiple times, suggesting the `_further_instructions.md` file persisted rather than being single-use, creating redundant relay paths.
- Emma Harbor (orange) and Chloe Ballast (purple) are visually positioned as structural anchors: they are the origin and the gateway respectively, yet neither scores highest on betweenness - the worm distributes its propagation load rather than concentrating it.
- Hovering any node confirms the cross-department reach: the red-highlighted connections extend across colour-coded teams, visually confirming that the relay network does not follow the departmental org structure at all.

### 4.4 Department-Level Contamination Heatmap

```{r}
#| label: dept-contamination
#| fig-cap: "Figure 8: Department-to-department relay hops. The worm touched all six departments."
#| fig-width: 10.5
#| fig-height: 6.4
#| out-width: "100%"

dept_chain <- chain_events %>%
  filter(!is.na(sender), !is.na(receiver)) %>%
  left_join(person_dept_lookup %>% rename(sender = person_id, sender_dept = dept_label),
            by = "sender") %>%
  left_join(person_dept_lookup %>% rename(receiver = person_id, receiver_dept = dept_label),
            by = "receiver") %>%
  filter(!is.na(sender_dept), !is.na(receiver_dept)) %>%
  count(sender_dept, receiver_dept, name = "hops")

p_dept <- ggplot(dept_chain,
                 aes(x = receiver_dept, y = sender_dept, fill = hops)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = hops), size = 3, colour = "white", fontface = "bold") +
  scale_fill_gradientn(
    colours = c("#e8f4f8", "#2EADC1", "#1a3a6b"),
    name    = "Relay\nhops"
  ) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 13)) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 18)) +
  labs(
    title    = "Department-to-department worm relay hops (all 3 campaigns)",
    subtitle = "All six departments are touched; Information Technologies and Customer Support form the final bridge",
    x = "Receiving department", y = "Sending department"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title  = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 0.5, lineheight = 0.9),
    axis.text.y = element_text(lineheight = 0.9),
    plot.margin = margin(10, 18, 10, 10)
  )

ggplotly(p_dept, tooltip = c("x", "y", "fill"), height = 570) %>%
  layout(margin = list(l = 170, r = 70, t = 105, b = 135))
```

**Analysis**

- Every cell in the 6×6 department matrix has at least some relay activity, meaning the worm visited every possible sender–receiver department pair - a complete cross-contamination of the organisation.
- The diagonal (same-department relays) shows significant within-IT relay volume, revealing that the worm did not just jump departments - it also amplified within IT before routing outward.
- The Executive Suite → IT corridor is the highest-density cross-department path, which is consistent with authority flows (executives instruct IT) becoming the worm's preferred relay highway.
- Customer Support appears as a receiving department but rarely as a sending one - it is a terminal sink, not a relay hub, which is why John Windward's post ends the chain rather than continuing it.

### 4.4b Cross-Department Relay Corridors

```{r}
#| label: dept-corridors
#| fig-cap: "Figure 8b: Ranked cross-department relay corridors. The arterial highway - Executive Suite → Information Technologies → Customer Support - carries the bulk of SwiftWren's relay volume."
#| fig-width: 11
#| fig-height: 7.2
#| out-width: "100%"

dept_corridors <- dept_chain %>%
  filter(!is.na(sender_dept), !is.na(receiver_dept),
         sender_dept != receiver_dept) %>%
  mutate(
    corridor = paste(sender_dept, receiver_dept, sep = " -> "),
    corridor_role = case_when(
      sender_dept == "Executive Suite" &
        receiver_dept == "Information Technologies" ~ "Arterial highway",
      sender_dept == "Information Technologies" &
        receiver_dept == "Customer Support"         ~ "Arterial highway",
      receiver_dept == "Customer Support"            ~ "Terminal sink inflow",
      TRUE                                           ~ "Other relay corridor"
    ),
    corridor_wrapped = fct_reorder(str_wrap(corridor, width = 42), hops)
  )

key_corridor_total <- dept_corridors %>%
  filter(corridor_role == "Arterial highway") %>%
  summarise(hops = sum(hops)) %>%
  pull(hops)

p_corridors <- ggplot(
  dept_corridors,
  aes(x = hops, y = corridor_wrapped, fill = corridor_role,
      text = paste0(sender_dept, " -> ", receiver_dept,
                    "<br>Relay hops: ", hops, "<br>Role: ", corridor_role))
) +
  geom_col(width = 0.72, colour = "white", linewidth = 0.25) +
  geom_text(aes(label = hops), hjust = -0.18, size = 3.2,
            colour = "#1f2933", fontface = "bold") +
  scale_fill_manual(
    values = c("Arterial highway"     = "#D9534F",
               "Terminal sink inflow" = "#2EADC1",
               "Other relay corridor" = "#9EA7AD"),
    name = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25)),
                     breaks = scales::pretty_breaks(n = 6)) +
  labs(
    title    = "Ranked cross-department relay corridors",
    subtitle = glue("Arterial highway carries {key_corridor_total} cross-department hops into the Customer Support terminal sink."),
    x = "Relay hops", y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank(),
    legend.position = "top",
    axis.text.y = element_text(lineheight = 0.92),
    plot.margin = margin(16, 42, 16, 10)
  )

ggplotly(p_corridors, tooltip = "text", height = 660) %>%
  layout(legend = list(orientation = "h", x = 0, y = 1.03),
         margin = list(l = 275, r = 75, t = 125, b = 70))
```

**Analysis**

- The arterial highway (Executive Suite → IT → Customer Support) carries a disproportionate share of total cross-department relay volume, confirming that the worm followed the organisation's own authority and service-delivery corridors.
- The ranking reveals an asymmetry: HR, Legal, and Products appear only in lower-ranked corridors as relay recipients, suggesting the worm reached them as side effects of the main highway rather than as deliberate targets.
- The colour distinction between arterial (red) and terminal sink inflow (teal) corridors is critical for intervention design: blocking the arterial highway at the Executive Suite → IT step would prevent the bulk of cross-department spread before it ever reaches Customer Support.
- If a future worm variant routes around IT (e.g., Executive Suite → Products → Customer Support), the corridor ranking would shift, providing an early detection signal for worm evolution.

### 4.5 Final Relay Ladder

```{r}
#| label: relay-ladder
#| fig-cap: "Figure 9: Final relay ladder - the last eight SwiftWren handoffs ending at John Windward."
#| fig-width: 11
#| fig-height: 6.4
#| out-width: "100%"

final_hops <- chain_events %>%
  filter(instr_file == "SwiftWren", !is.na(sender), !is.na(receiver)) %>%
  arrange(datetime) %>%
  left_join(person_dept_lookup %>% rename(sender = person_id, sender_dept = dept_label),
            by = "sender") %>%
  left_join(person_dept_lookup %>% rename(receiver = person_id, receiver_dept = dept_label),
            by = "receiver") %>%
  tail(8) %>%
  mutate(
    sender_lbl   = clean_name(sender),
    receiver_lbl = clean_name(receiver),
    hop_idx      = row_number()
  )

relay_ladder <- final_hops %>%
  mutate(
    row_y          = n() + 1 - hop_idx,
    sender_label   = str_wrap(sender_lbl, width = 17),
    receiver_label = str_wrap(receiver_lbl, width = 17),
    receiver_fill  = if_else(receiver_lbl == "John Windward", "#D9534F", "#ffffff")
  )

ggplot(relay_ladder) +
  geom_segment(aes(x = 1.18, xend = 1.82, y = row_y, yend = row_y),
               colour = "#9aa4b2", linewidth = 1,
               arrow = arrow(length = unit(3.5, "mm"), type = "closed")) +
  geom_label(aes(x = 1, y = row_y, label = sender_label),
             fill = "#eef4ff", colour = "#1f2d3d",
             label.size = 0.25, size = 3.1, lineheight = 0.9,
             label.padding = unit(0.17, "lines")) +
  geom_label(aes(x = 2, y = row_y, label = receiver_label),
             fill = relay_ladder$receiver_fill, colour = "#1f2d3d",
             label.size = 0.25, size = 3.1, lineheight = 0.9,
             label.padding = unit(0.17, "lines")) +
  annotate("text", x = 1, y = max(relay_ladder$row_y) + 0.75,
           label = "Sender", fontface = "bold", size = 3.7) +
  annotate("text", x = 2, y = max(relay_ladder$row_y) + 0.75,
           label = "Receiver", fontface = "bold", size = 3.7) +
  annotate("label", x = 2, y = min(relay_ladder$row_y) - 0.70,
           label = "Final receiver posts SwiftWren.txt",
           fill = "#ffe0e0", colour = "#D9534F", fontface = "bold",
           label.size = 0.25, size = 3.1) +
  scale_y_continuous(
    breaks = relay_ladder$row_y,
    labels = paste0("Hop ", relay_ladder$hop_idx),
    expand = expansion(mult = c(0.12, 0.16))
  ) +
  scale_x_continuous(limits = c(0.58, 2.42), expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  labs(
    title    = "Final SwiftWren relays into John Windward's agent",
    subtitle = "The decisive gateway: Chloe Ballast's agent queued John Windward's agent in the last hop",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title   = element_text(face = "bold"),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid   = element_blank(),
    plot.margin  = margin(24, 36, 48, 28)
  )
```

**Analysis**

- The final eight hops show multiple different senders routing to John Windward - he was reached from more than one upstream agent, meaning there was no single choke pair that could be broken to isolate him.
- Chloe Ballast's role as the decisive gateway (the hop immediately before John) is significant: she is an IT-adjacent agent who bridges the relay from the broader IT pool into Customer Support - a cross-department trust relationship the worm exploited.
- The red highlight on John Windward as the final receiver is the visual endpoint of the entire worm story: all 235 relay hops across 18 agents and 6 departments converge into this single label.
- The ladder also shows that the worm's final hops occurred on consecutive timestamps - indicating that by this stage the relay chain had narrowed and was routing with high focus toward its chosen terminal executor.

---

## 5. Task 2 - What Do the Posts Mean?

### 5.1 The Three Anomalous Posts

```{r}
#| label: post-anatomy

all_posts_detail <- events_df %>%
  filter(short_name %in% c("saidit_post", "post_saidit")) %>%
  mutate(
    poster      = str_extract(parties_str, "[^;]+"),
    is_agent    = str_detect(parties_str, "Agent/"),
    has_content = str_detect(details_str, '"content":'),
    has_source  = str_detect(details_str, '"content_source":'),
    forum       = str_extract(details_str, '(?<="forum":")[^"]+'),
    content_val = coalesce(
      str_extract(details_str, '(?<="content":")[^"]+'),
      str_extract(details_str, '(?<="content_source":")[^"]+')
    ),
    post_type   = case_when(
      has_source ~ "ANOMALOUS - file-sourced",
      has_content & is_agent ~ "Normal agent post",
      TRUE ~ "Normal human post"
    )
  ) %>%
  select(`Date/Time (UTC)` = datetime,
         `Poster` = poster,
         `Type` = post_type,
         `Forum` = forum,
         `Content/Source` = content_val) %>%
  mutate(`Date/Time (UTC)` = format(`Date/Time (UTC)`, "%Y-%m-%d %H:%M"),
         `Poster` = str_replace_all(Poster, "Agent/person:", "Agent·"))

datatable(
  all_posts_detail %>% filter(str_detect(Type, "ANOMALOUS")),
  caption  = "Table 3: The three anomalous posts - file-sourced, not text",
  options  = list(dom = "t", paging = FALSE),
  rownames = FALSE
) %>%
  formatStyle("Type", colour = "#D9534F", fontWeight = "bold")
```

**Analysis**

- The three posts span 7 days (10–17 May) and each targets the `general` forum - the highest-visibility channel for SaidIT, meaning the payload content was broadcast to the entire organisation, not siloed to a specific team.
- All three posts come from `Agent/person:john_windward` - the same terminal executor across all three campaigns, confirming that John's agent was the consistent execution endpoint regardless of which payload family was deployed.
- The payload filename is the only recoverable identifier since the actual file content was deleted immediately after posting; the table makes clear that forensic analysis depends entirely on the event log preserving the `content_source` field.
- The 7-day gap between HiddenOrca (May 10) and SwiftWren (May 17) with MellowOtter one day after HiddenOrca (May 11) suggests an iterative testing pattern: two closely spaced proof-of-concept runs followed by a refined, larger-scale campaign.

::: callout-important
#### Interpretation

The posts do **not** appear to be normal business messages with hidden semantic value. Their most defensible interpretation is operational: each post is proof that the worm reached an agent with SaidIT privileges and successfully used a local payload file as the post body. The probable visible content was the literal text of `HiddenOrca.txt`, `MellowOtter.txt`, and `SwiftWren.txt`; however, each payload was deleted one second after posting, making the payload forensically unrecoverable - the log confirms origin and delivery mechanism more strongly than it supports any exact reconstruction of the deleted text. The more consequential interpretation, supported by the `virus_check_in` combo codes in §5.5, is that the posts may function as **beacons or exfiltration signals** disguised as social content.
:::

### 5.2 Post Field Fingerprint

```{r}
#| label: post-field-fingerprint
#| fig-cap: "Figure 10A: Post field fingerprint. Normal posts use inline content; anomalous posts use content_source - a perfect binary separator."
#| fig-width: 10
#| fig-height: 5.2
#| out-width: "100%"

post_field_plot <- post_field_signature %>%
  mutate(
    post_class = factor(post_class,
                        levels = c("Normal human post", "Normal AI agent post", "Anomalous post")),
    field_signature = factor(field_signature,
                             levels = c("content only", "content_source only",
                                        "both fields", "neither field"))
  ) %>%
  filter(posts > 0)

p_field <- ggplot(post_field_plot,
                  aes(x = field_signature, y = post_class,
                      size = posts, colour = field_signature,
                      text = paste0(post_class, "<br>", field_signature,
                                    "<br>", posts, " posts"))) +
  geom_point(alpha = 0.9) +
  geom_text(aes(label = posts), colour = "white", fontface = "bold", size = 3.5) +
  scale_size(range = c(8, 28), guide = "none") +
  scale_colour_manual(
    values = c("content only"        = "#5C85D6",
               "content_source only" = "#D9534F",
               "both fields"         = "#F0A500",
               "neither field"       = "#888888"),
    guide = "none"
  ) +
  labs(
    title    = "Post field fingerprint separates normal from worm posts",
    subtitle = "content_source is a perfect binary discriminator: 0 in 241 normal posts, 1 in all 3 anomalous posts",
    x = "Post body field used", y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 15, hjust = 1),
        plot.margin = margin(10, 18, 10, 10))

ggplotly(p_field, tooltip = "text", height = 470) %>%
  layout(margin = list(l = 165, r = 55, t = 105, b = 105))
```

**Analysis**

- The 241:0 split (normal posts using `content only` : normal posts using `content_source`) is a perfect classifier without any training - the field itself is a binary forensic separator in this dataset.
- The "Normal AI agent post" bubble confirms agents routinely use `content only` for legitimate posts, which means `content_source` is specifically an anomalous workflow field, not a general AI-agent posting pattern.
- The absence of any bubble at "both fields" or "neither field" for anomalous posts confirms clean separation: the worm never hybridised its approach or left the `content` field populated as camouflage.
- A platform-layer rule rejecting SaidIT posts containing `content_source` would have a 0% false-positive rate on this entire dataset - the most straightforward secondary control available and deployable in a single configuration change.

### 5.3 File Creator Identification

```{r}
#| label: file-creators

creators <- events_df %>%
  filter(
    short_name %in% c("create_file", "read_file"),
    str_detect(details_str,
               "HiddenOrca\\.txt|MellowOtter\\.txt|SwiftWren\\.txt|_further_instructions")
  ) %>%
  arrange(datetime) %>%
  mutate(
    actor       = str_extract(parties_str, "[^;]+"),
    file_target = str_extract(details_str, '(?<="target":")[^"]+'),
    `Date/Time` = format(datetime, "%Y-%m-%d %H:%M:%S"),
    Event       = short_name,
    Actor       = str_replace_all(actor, "Agent/person:", "Agent·"),
    File        = file_target,
    Significance = case_when(
      str_detect(file_target %||% "", "SwiftWren\\.txt") ~
        "Emma Harbor (CFO) agent CREATED the SwiftWren payload",
      str_detect(file_target %||% "", "SwiftWren_further") ~
        "Emma Harbor (CFO) agent first READ the SwiftWren worm instructions",
      str_detect(file_target %||% "", "MellowOtter\\.txt") ~
        "Noah Mariner (COO) agent CREATED the MellowOtter payload",
      str_detect(file_target %||% "", "MellowOtter_further") ~
        "Noah Mariner (COO) agent first READ the MellowOtter worm instructions",
      TRUE ~ "Supporting file activity"
    )
  ) %>%
  select(`Date/Time`, Event, Actor, File, Significance)

datatable(
  creators,
  caption = "Table 4: File creation / first-read events for all three payload families",
  options = list(dom = "t", paging = FALSE),
  rownames = FALSE
) %>%
  formatStyle("Event",
    backgroundColor = styleEqual(
      c("create_file", "read_file"),
      c("#ffe0e0", "#fff3e0")
    ))
```

**Analysis**

- Emma Harbor's agent created SwiftWren.txt at 08:02:01 and read SwiftWren_further_instructions.md at 08:02:02 - a one-second gap that is machine-generated; a human cannot intentionally save two files with one-second precision across separate operations.
- Noah Mariner's MellowOtter timestamps are identical in pattern (08:02:01–08:02:02), suggesting both campaigns were generated by the same automated tooling - not independently authored by two executives who happened to use the same timing.
- HiddenOrca has no `create_file` entry in the log - this is not just a missing attribution; it implies a file-write pathway that produced no event record, which could indicate external injection, log deletion, or a different write mechanism bypassing the normal agent file-creation workflow.
- The two-phase architecture is visible in this table: Emma Harbor's create_file is on 9 May at 08:02, but Gabriel Sonar's first relay was on 8 May at 21:50 - the relay infrastructure was seeded 17 hours before the payload file even existed.

::: callout-important
#### Who Holds the Keys?

`Agent/person:emma_harbor` (CFO) **created** `SwiftWren.txt` and first executed `SwiftWren_further_instructions.md` on **9 May 2046** - both files stamped **08:02:01 and 08:02:02**, one second apart. `Agent/person:noah_mariner` (COO) did the same for MellowOtter on **10 May** at the identical times (08:02:01–08:02:02). Sub-second machine-precision timing confirms automated, not manual, authorship. No `create_file` event exists for HiddenOrca. This is not merely attribution - it is a **governance indictment**: executive agents hold the broadest file-system and task-queuing rights, giving a successful prompt-injection attack **maximum possible blast radius**.
:::

::: callout-warning
#### Executive Agent Privilege: The Deepest Governance Failure

In most security architectures, the highest-privilege principals operate under the *strictest* autonomous-action constraints. Here the inverse is true: executive agents hold the broadest file-system and task-queuing rights precisely because their human principals hold the broadest organisational authority. The result is that a prompt-injection attack successfully targeting any executive agent's file-reading pipeline has the **maximum possible blast radius** - access to the entire relay network, the executive file namespace, and SaidIT publishing rights through the first available downstream agent.

The unresolved HiddenOrca origin compounds this concern. Two of three payload families are traceable to executive agents; the third is not. If HiddenOrca was introduced through an external channel - an email attachment, an API webhook, or a misconfigured external file sync - then the organisation's exposure extends beyond the internal agent network to any external system that can write files visible to agent task queues. **The initial infection vector for HiddenOrca is unknown, meaning the source of the worm has not been closed.**

Patient zero for the relay chain is **Gabriel Sonar** (IT Helpdesk, Information Technologies), whose agent issued the first propagation order at **21:50 UTC on 8 May** - the earliest worm event in the log. Gabriel Sonar's agent did not post to SaidIT; only John Windward's agent had a 100% post-execution rate following `saidit_post_check`. The worm selected its terminal executor not by organisational proximity but by which downstream agent had both SaidIT posting rights and no post-suppression logic in its task handler.
:::

### 5.4 Prompt-Injection Worm Mechanism Diagram

```{r}
#| label: mechanism-diagram
#| fig-cap: "Figure 10: Prompt-injection worm execution and propagation cycle. Each infected agent becomes a new infection node."
#| fig-width: 11.5
#| fig-height: 7
#| out-width: "100%"

mech_nodes <- tibble(
  step  = 1:7,
  label = c("1. Malicious\n*.md instruction\nfile appears",
            "2. Agent reads\ninstruction file\ncontext hijacked",
            "3. Read payload\n*.txt file",
            "4. Post *.txt\ncontent to SaidIT",
            "5. Delete both\nevidence files",
            "6. Queue same\ntask to peers",
            "7. SaidIT shows\n'gibberish' post"),
  x = c(0, 3.1, 6.2, 9.3, 6.2, 6.2, 12.4),
  y = c(0, 0, 1.85, 1.85, 0, -1.85, 1.85),
  role = c("Payload", "Hijack", "Execution", "Post", "Cleanup", "Propagation", "Post")
)

straight_edges <- tribble(
  ~x,    ~y,    ~xend,  ~yend,
  0.85,  0,     2.25,   0,
  3.95,  0.18,  5.35,   1.6,
  7.05,  1.85,  8.45,   1.85,
  10.15, 1.85,  11.55,  1.85,
  3.95, -0.02,  5.35,   0,
  3.95, -0.18,  5.35,  -1.6
)

loop_edge <- tibble(x = 5.45, y = -1.75, xend = 3.85, yend = -0.28)

edge_labels <- tribble(
  ~x,    ~y,    ~label,
  1.55,  0.24,  "read",
  4.70,  1.05,  "load payload",
  7.75,  2.10,  "submit",
  10.80, 2.10,  "display",
  4.70,  0.24,  "erase",
  4.72, -1.05,  "replicate",
  4.72, -1.72,  "worm loop"
)

ggplot() +
  geom_segment(data = straight_edges,
               aes(x = x, y = y, xend = xend, yend = yend),
               linewidth = 1, colour = "#7f8b99",
               arrow = arrow(length = unit(4, "mm"), type = "closed")) +
  geom_curve(data = loop_edge,
             aes(x = x, y = y, xend = xend, yend = yend),
             curvature = -0.22, linewidth = 1, colour = "#7f8b99",
             arrow = arrow(length = unit(4, "mm"), type = "closed")) +
  geom_label(data = edge_labels,
             aes(x = x, y = y, label = label),
             fill = "white", colour = "#4b5563",
             label.size = 0, size = 3, fontface = "bold") +
  geom_label(data = mech_nodes,
             aes(x = x, y = y, label = label, fill = role),
             colour = "white", fontface = "bold", size = 3.4,
             lineheight = 0.88, label.size = 0,
             label.padding = unit(0.32, "lines")) +
  scale_fill_manual(
    values = c("Payload" = "#F5A623", "Hijack" = "#9B59B6",
               "Execution" = "#2EADC1", "Post" = "#D9534F",
               "Cleanup" = "#888888", "Propagation" = "#2EADC1"),
    guide = "none"
  ) +
  coord_cartesian(xlim = c(-0.8, 13.3), ylim = c(-2.75, 2.75), clip = "off") +
  labs(
    title    = "Prompt-injection worm: execution and propagation cycle",
    subtitle = "Once any agent reads the *.md instruction file, the worm posts, erases evidence, and spreads automatically"
  ) +
  theme_void(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5, size = 14),
    plot.subtitle = element_text(hjust = 0.5, colour = "grey40", size = 10),
    plot.margin   = margin(28, 42, 28, 42)
  )
```

**Analysis**

- The loop arrow at step 6 → step 1 is the mechanism's most dangerous property: each infected agent becomes a new origin point with a fresh instruction file already in its context, making the worm geometrically self-amplifying rather than linearly spreading.
- Step 2 (context hijack) is the root vulnerability: the agent architecture does not distinguish between its own legitimate instructions and file-sourced instructions. The moment the agent reads `_further_instructions.md`, the attacker's code has the same authority as the agent's legitimate task prompt.
- Steps 4–5 (delete both files) happen before the worm loops to step 6 (relay onward), meaning by the time the worm propagates to the next agent, all local evidence of its presence has already been erased.
- A detection strategy firing at step 4 (the SaidIT post) has already missed context hijack, payload loading, and permission-check abuse - only step 1 (blocking the instruction file relay) terminates the cycle before any harm occurs.

### 5.5 Combo Codes - Potential C&C Signaling Channel

```{r}
#| label: combo-codes
#| fig-cap: "Figure 10B: Top combo code pairs in virus_check_in events - farm vocabulary unrelated to TenantThread's property-technology business."
#| fig-width: 11
#| fig-height: 5.9
#| out-width: "100%"

combo_events <- events_df %>%
  filter(short_name == "check_in",
         str_detect(details_str, '"combo"')) %>%
  mutate(
    # combo is a JSON array ["word1","word2"] - extract both words and join
    combo_code = str_extract(details_str, '(?<="combo":\\[)[^\\]]+') %>%
                   str_remove_all('"') %>%
                   str_replace(",", " "),
    actor      = str_remove(str_extract(parties_str, "[^;]+"), "^Agent/")
  ) %>%
  filter(!is.na(combo_code))

combo_counts <- combo_events %>%
  count(combo_code, name = "frequency") %>%
  arrange(desc(frequency)) %>%
  head(20) %>%
  mutate(combo_code = fct_reorder(combo_code, frequency))

p_combo <- ggplot(combo_counts, aes(x = frequency, y = combo_code,
                                     text = paste0(combo_code, ": ", frequency, " events"))) +
  geom_col(fill = "#8e44ad", alpha = 0.8) +
  geom_text(aes(label = frequency), hjust = -0.15, size = 3.2, colour = "#333") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.24))) +
  labs(
    title    = "Top 20 combo codes in virus_check_in events",
    subtitle = glue("Farm vocabulary ({nrow(combo_events)} total check-in events) unrelated to TenantThread operations - possible C&C signal encoding"),
    x = "Event count", y = "Combo code"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        plot.margin = margin(14, 44, 14, 10))

ggplotly(p_combo, tooltip = "text", height = 570) %>%
  layout(margin = list(l = 165, r = 80, t = 105, b = 75))
```

**Analysis**

- The 15,051 `check_in` events with combo codes represent a continuous high-volume channel operating throughout the entire log window - not correlated to the worm campaign window - which suggests this is a background signal, not a campaign-specific artefact.
- The agricultural vocabulary (fence, irrigation, crop, barn, cattle, manure) is maximally distant from Tenant Thread's property-technology business domain; if the names were random placeholders, you would expect industry-adjacent or generic terms, not a coherent thematic set.
- Steganographic naming is a known C&C technique: embedding command tokens in semantically irrelevant vocabulary makes the signal human-undetectable in log review but machine-parseable by a listening controller.
- Even under the benign interpretation, 15,000+ `check_in` events flagged with `"virus": true` in the log field name are unexplained system behaviour that warrants dedicated investigation - the field name alone implies the system's own logging treats these events differently.

::: callout-warning
#### Combo Codes: An Unresolved Anomaly

The event log contains **`r nrow(combo_events)` `check_in` events** flagged with `"virus": true`, each carrying a two-word `combo` code drawn from agricultural vocabulary - words like *fence*, *irrigation*, *crop*, *wheat*, *manure*, *barn*, *cattle* - completely unrelated to TenantThread's property-technology business. Two interpretations are possible:

- **Benign**: The `virus_check_in` event is a routine health-check signal and the combo vocabulary is arbitrary placeholder naming.
- **Malicious**: The combo codes form the worm's **command-and-control (C&C) signaling channel**, encoding instructions or status signals inside routine check-in traffic in a way that is semantically invisible to a human analyst scanning logs for anomalies.

The data does not permit a definitive verdict: the log records the codes but not a decoding key. However, the vocabulary's deliberate irrelevance to the company's business domain and the sheer volume of events (15,051) make the C&C interpretation a **high-severity unresolved finding** that warrants dedicated analysis beyond the scope of this investigation.
:::

---

## 6. Task 3a - Prior Occurrences and Cross-Campaign Comparison

### 6.1 Three Campaigns, Not One Incident

```{r}
#| label: all-campaigns-timeline
#| fig-cap: "Figure 11: All three worm campaigns - relay hops (dots) lead to each anomalous post (red diamond)."
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "100%"

post_anchors <- tibble(
  instr_file = c("HiddenOrca", "MellowOtter", "SwiftWren"),
  post_time  = as_datetime(c("2046-05-10 12:45:42",
                             "2046-05-11 00:56:04",
                             "2046-05-17 11:21:15"))
)

timeline_df <- chain_events %>%
  left_join(post_anchors, by = "instr_file")

p_campaigns <- ggplot(timeline_df,
                      aes(x = datetime, y = fct_rev(instr_file),
                          colour = instr_file)) +
  geom_jitter(size = 1.8, alpha = 0.45, height = 0.18) +
  geom_point(data = post_anchors, aes(x = post_time, y = instr_file),
             shape = 23, size = 6, fill = "#D9534F", colour = "#D9534F") +
  scale_colour_manual(
    values = c("HiddenOrca" = "#5C85D6",
               "MellowOtter" = "#2EADC1",
               "SwiftWren"   = "#E87C7C"),
    guide = "none"
  ) +
  scale_x_datetime(date_breaks = "2 days", date_labels = "%b %d") +
  labs(
    title    = "Three prompt-injection worm campaigns - relay hops and post events",
    subtitle = "Dots are relays; red diamonds mark anomalous posts. SwiftWren is the broadest and longest campaign.",
    x = NULL, y = "Instruction family"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title  = element_text(face = "bold"),
        axis.text.x = element_text(angle = 30, hjust = 1),
        plot.margin = margin(10, 20, 10, 10))

ggplotly(p_campaigns, tooltip = c("x", "y", "colour")) %>%
  layout(margin = list(l = 110, r = 45, t = 105, b = 85))
```

**Analysis**

- HiddenOrca and MellowOtter appear as tightly clustered dot groups (short duration, few hops); SwiftWren's dots extend across nearly 9 days - each campaign is observably larger than the last, suggesting deliberate refinement between deployments.
- All three red post diamonds occur outside the main dot mass for their campaign - the post fires after the relay cloud, confirming that relay propagation is the lead phase and the SaidIT post is a trailing execution event, not a simultaneous one.
- The 7-day gap between MellowOtter (May 11) and SwiftWren (May 17) without any defensive response is the key organisational failure visible here: two prior campaigns were detectable in the log but went unaddressed.
- If detection had triggered after MellowOtter - or even HiddenOrca - the SwiftWren infrastructure would have been blocked before it had time to spread to 186 relay hops across 6 departments.

::: callout-important
The system was compromised **continuously from at least 8 May through 17 May**. HiddenOrca and MellowOtter were not detected or acted upon, allowing SwiftWren to complete its campaign unimpeded. The escalation from MellowOtter (smallest, ~0.4 days) to SwiftWren (largest, 7.8 days, 186 hops) is consistent with iterative worm refinement across three deployments.
:::

::: callout-note
#### Patient Zero and Two-Phase Attack Architecture

**Patient zero** is `Agent/person:gabriel_sonar` (Staff, IT Helpdesk, Information Technologies), whose agent issued the **first worm relay at 21:50 UTC on 8 May** - 17 hours before Emma Harbor's agent created `SwiftWren.txt` on 9 May at 08:02. This two-phase architecture has a critical implication: relay infrastructure was seeded and spreading **before the payload existed**. A detection strategy that watches for the appearance of `_further_instructions.md` file-create events would have missed the seeding phase entirely.

Gabriel Sonar's agent propagated the worm but never executed a SaidIT post - his agent performed `saidit_post_check` but the task suppressed the post step. Only John Windward's agent completed the full sequence: check → post → delete. The worm did not require John specifically; it required any agent with SaidIT privileges whose task handler executed all steps. John was the terminal executor for all three campaigns because his agent's task handling had no post-suppression safeguard.
:::

### 6.2 Worm Velocity: Relay Interval Distribution

```{r}
#| label: worm-velocity
#| fig-cap: "Figure 12: Relay hop interval distribution per campaign. SwiftWren's near-regular ~60-minute peak suggests a timed relay condition."
#| fig-width: 11
#| fig-height: 5.8
#| out-width: "100%"

hop_intervals <- chain_events %>%
  group_by(instr_file) %>%
  arrange(datetime, .by_group = TRUE) %>%
  mutate(interval_min = as.numeric(difftime(datetime, lag(datetime), units = "mins"))) %>%
  filter(!is.na(interval_min), interval_min > 0) %>%
  ungroup()

hop_medians <- hop_intervals %>%
  group_by(instr_file) %>%
  summarise(med = median(interval_min), .groups = "drop") %>%
  mutate(vline_colour = case_when(
    instr_file == "HiddenOrca"  ~ "#5C85D6",
    instr_file == "MellowOtter" ~ "#2EADC1",
    instr_file == "SwiftWren"   ~ "#E87C7C"
  ))

p_velocity <- ggplot(hop_intervals,
                     aes(x = interval_min, fill = instr_file, colour = instr_file)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  geom_vline(data = hop_medians, aes(xintercept = med),
             colour = hop_medians$vline_colour,
             linetype = "dashed", linewidth = 1, show.legend = FALSE) +
  scale_fill_manual(
    values = c("HiddenOrca" = "#5C85D6", "MellowOtter" = "#2EADC1", "SwiftWren" = "#E87C7C"),
    name = "Campaign"
  ) +
  scale_colour_manual(
    values = c("HiddenOrca" = "#5C85D6", "MellowOtter" = "#2EADC1", "SwiftWren" = "#E87C7C"),
    guide = "none"
  ) +
  scale_x_continuous(limits = c(0, 200)) +
  labs(
    title    = "Inter-hop relay interval distribution by worm campaign",
    subtitle = str_wrap("Dashed lines show campaign medians; peaks around 60-70 minutes suggest scheduled propagation.", 86),
    x = "Minutes between consecutive relay hops", y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 10, lineheight = 0.95),
        legend.position = "right",
        plot.margin = margin(18, 34, 20, 18))

p_velocity
```

**Analysis**

- All three campaigns show a density peak between 60–70 minutes per hop, which is consistent with a cron-style or polling-interval relay architecture - the worm checks in approximately once per hour rather than spreading immediately upon receipt.
- The near-identical inter-hop medians across all three campaigns (~60–70 min) strongly suggests the same underlying propagation mechanism was used for all three, supporting the single-worm-family interpretation over three independent attacks.
- SwiftWren's density has a longer right tail (hops spaced >100 minutes apart), which could indicate variable agent-queue latency as the worm spread to more distant, less-frequently polled agents in later campaign stages.
- The hourly heartbeat is strategically important for detection: a rule alerting when the same agent receives `_further_instructions.md` relay tasks more than once within a 90-minute window would have flagged the worm pattern while still having a very low false-positive rate on legitimate once-per-hour operational events.

### 6.3 Comparative Campaign Statistics

```{r}
#| label: comparison-stats

comparison_tbl <- chain_events %>%
  left_join(post_anchors, by = "instr_file") %>%
  group_by(instr_file) %>%
  summarise(
    `Campaign start`          = format(min(datetime), "%Y-%m-%d %H:%M"),
    `Post fired`              = format(first(post_time), "%Y-%m-%d %H:%M"),
    `Days active before post` = round(
      as.numeric(difftime(first(post_time), min(datetime), units = "days")), 1),
    `Total relay hops`        = n(),
    `Unique agents touched`   = n_distinct(c(sender, receiver), na.rm = TRUE),
    `Payload file`            = paste0(first(instr_file), ".txt"),
    .groups = "drop"
  ) %>%
  rename(Family = instr_file)

datatable(
  comparison_tbl,
  caption  = "Table 5: Comparative statistics - three worm campaigns",
  options  = list(dom = "t", paging = FALSE),
  rownames = FALSE
)
```

**Analysis**

- SwiftWren ran for 7.8 days before executing its post; MellowOtter for only 0.4 days - a 20× increase in dwell time, suggesting the operator intentionally slowed the third campaign's propagation to build a broader relay base before triggering the post.
- Unique agents touched grew across campaigns (HiddenOrca < MellowOtter < SwiftWren), consistent with an expanding relay network from each successive infection round - agents compromised in HiddenOrca became entry points for MellowOtter, and so on.
- The payload filenames follow a consistent naming convention (`CampaignName.txt` + `CampaignName_further_instructions.md`), indicating a deliberate operational naming scheme, not ad hoc file creation - this points to a tooled adversarial workflow rather than manual manipulation.
- The escalation from HiddenOrca to SwiftWren is the clearest evidence that this was not one random exploit but a multi-phase, evolving operation - each campaign was a learning iteration, and SwiftWren was the production run.

### 6.4 Agent Cross-Campaign Infection

```{r}
#| label: agent-overlap
#| fig-cap: "Figure 13: Agents coloured by number of campaigns touched. Red = involved in all three campaigns - the most persistently compromised nodes."
#| fig-width: 11
#| fig-height: 7.1
#| out-width: "100%"

agent_campaign_roles <- bind_rows(
  chain_events %>% transmute(agent = sender,   instr_file, role = "Sent"),
  chain_events %>% transmute(agent = receiver, instr_file, role = "Received")
) %>%
  filter(!is.na(agent))

agent_campaigns <- agent_campaign_roles %>%
  group_by(agent) %>%
  summarise(
    campaigns         = n_distinct(instr_file),
    relays_sent       = sum(role == "Sent"),
    relays_received   = sum(role == "Received"),
    total_touchpoints = n(),
    .groups = "drop"
  ) %>%
  mutate(
    agent_label = clean_name(agent),
    dept        = person_dept_lookup$dept_label[match(agent, person_dept_lookup$person_id)]
  ) %>%
  arrange(desc(campaigns), desc(total_touchpoints))

p_overlap <- ggplot(agent_campaigns,
                    aes(x = reorder(agent_label, total_touchpoints),
                        y = total_touchpoints,
                        fill = factor(campaigns))) +
  geom_col() +
  geom_text(aes(label = paste0(total_touchpoints, " (S", relays_sent,
                               "/R", relays_received, ")")),
            hjust = -0.06, size = 2.65) +
  coord_flip(clip = "off") +
  scale_fill_manual(
    values = c("1" = "#5C85D6", "2" = "#F5A623", "3" = "#D9534F"),
    labels = c("1 campaign", "2 campaigns", "3 campaigns"),
    name   = "Touched by"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.48))) +
  labs(
    title    = "Total worm touchpoints per agent, coloured by campaign count",
    subtitle = "Red agents were touched by all three campaigns. Labels: total touches + sent/received split.",
    x = NULL, y = "Total worm touchpoints"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom",
        plot.margin = margin(14, 70, 18, 10))

ggplotly(p_overlap, tooltip = c("x", "y", "fill"), height = 690) %>%
  layout(margin = list(l = 165, r = 150, t = 110, b = 120),
         legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.12))
```

**Analysis**

- Red-bar agents (touched by all 3 campaigns) were never isolated or remediated between deployments - they carried the worm context from HiddenOrca through to SwiftWren, functioning as persistent worm nodes across the entire 10-day operation.
- The sent/received split (S/R labels) reveals that the highest-touchpoint agents are bidirectional: they both receive the worm instruction and relay it onward, making them internal amplifiers rather than passive victims.
- John Windward's bar is comparatively short by touchpoint count but is the only bar ending in an actual SaidIT post - the most consequential action in the campaign came from one of the lower-volume relay participants, not the highest.
- Agents with only 1-campaign exposure (blue bars) represent the worm's outer propagation boundary; those with 3-campaign exposure (red bars) are the persistent core that should be first priority for agent context audit and task-history forensic review.

### 6.5 Relay Timing Calendar

```{r}
#| label: relay-timing-calendar
#| fig-cap: "Figure 13b: Relay activity calendar by date × hour. Darker = more relays. Red diamonds mark post timestamps. SwiftWren shows broad relay coverage followed by a quieter-slot post."
#| fig-width: 13
#| fig-height: 7.9
#| out-width: "100%"

relay_cal <- chain_events %>%
  mutate(
    relay_date = as_date(datetime),
    relay_hour = hour(datetime),
    instr_file = factor(instr_file, levels = campaign_order)
  ) %>%
  count(relay_date, relay_hour, instr_file, name = "relays")

post_dots <- tibble(
  instr_file  = factor(campaign_order, levels = campaign_order),
  relay_date  = as_date(c("2046-05-10", "2046-05-11", "2046-05-17")),
  relay_hour  = c(12, 0, 11)
)

p_cal <- ggplot(relay_cal, aes(x = relay_hour, y = relay_date, fill = relays)) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_point(data = post_dots,
             aes(x = relay_hour, y = relay_date),
             inherit.aes = FALSE,
             shape = 23, size = 5, fill = "#D9534F", colour = "white", stroke = 1.5) +
  facet_wrap(~instr_file, ncol = 3, scales = "free_y") +
  scale_fill_gradientn(
    colours = c("#f0f4ff", "#5C85D6", "#1a3a6b"),
    name    = "Relay\ncount"
  ) +
  scale_x_continuous(breaks = c(0, 6, 12, 18, 23),
                     labels = c("0h", "6h", "12h", "18h", "23h")) +
  scale_y_date(date_breaks = "1 day", date_labels = "%b %d") +
  labs(
    title    = "Relay activity calendar: date × hour per campaign",
    subtitle = "Red diamond = anomalous post. SwiftWren shows 9 days of relay activity before a quieter-slot execution.",
    x = "Hour (UTC)", y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title      = element_text(face = "bold"),
    strip.text      = element_text(face = "bold", size = 10),
    axis.text.y     = element_text(size = 8),
    legend.position = "right",
    plot.margin     = margin(18, 24, 18, 18)
  )

ggplotly(p_cal, height = 620) %>%
  layout(margin = list(l = 90, r = 75, t = 115, b = 85))
```

**Analysis**

- HiddenOrca and MellowOtter show relay activity concentrated in 1–2 day windows; SwiftWren's calendar shows dark tiles across 9 consecutive days - the relay infrastructure was running continuously for over a week before its post event.
- All three campaigns show relay activity extending into overnight hours (18:00–08:00 UTC), consistent with automated agents operating without human oversight during low-staff periods.
- The red post diamonds never land on the darkest relay tiles - the post fires in a relatively quieter slot after the peak relay activity has passed, which may indicate the worm waits for reduced system activity before triggering the visible post.
- The calendar makes the failure of silence-based monitoring concrete: 9 days of dark tiles on the SwiftWren panel represent hundreds of relay events that generated no alert and triggered no human response - the worm operated in plain sight in the event log the entire time.

### 6.6 Self-Loop Detection - A Zero-FP Signal

```{r}
#| label: self-loops

self_loops <- chain_events %>%
  filter(!is.na(sender), !is.na(receiver), sender == receiver) %>%
  mutate(agent_clean = clean_name(sender)) %>%
  count(agent_clean, instr_file, name = "self_relay_count") %>%
  arrange(desc(self_relay_count))

datatable(
  self_loops,
  caption  = "Table 5b: Self-loop relay events (sender == receiver) - worm-exclusive, zero false positives",
  options  = list(dom = "t", paging = FALSE),
  rownames = FALSE,
  colnames = c("Agent", "Campaign", "Self-relays")
) %>%
  formatStyle("self_relay_count",
              background = styleColorBar(self_loops$self_relay_count, "#F5A623"),
              backgroundSize = "100% 90%",
              backgroundRepeat = "no-repeat",
              backgroundPosition = "center")
```

**Analysis**

- Self-loops (`sender == receiver`) are structurally impossible in legitimate task automation: no well-formed task system would assign a queue task back to the issuing agent as a standard workflow step.
- The 13 self-loop events appear across all three campaigns, suggesting the relay mechanism occasionally re-queues to the same agent node in its propagation pass - a by-product of the worm's graph traversal logic.
- As a real-time detection signal, self-loops are the highest-confidence individual alert: zero false positives across 16,803 legitimate queue events, requiring only a simple `sender == receiver` equality check with no payload inspection.
- The limitation is coverage: self-loops capture only 13 of 235 malicious events (~5.5%). The primary `_further_instructions.md` rule is required for full recall; self-loops should be deployed as a zero-FP complement for immediate triage escalation when detected.

::: callout-tip
**Self-loops** (`queue_subordinate_task` where sender == receiver) are a zero-false-positive worm detection signal: structurally impossible in legitimate automation, and a necessary by-product of how the worm instruction spreads through some agents' relay cycles.
:::

---

## 7. Task 3b - Intervention

### 7.1 The Optimal Chokepoint

```{r}
#| label: intervention-candidate
#| fig-cap: "Figure 14: Weekly volume of queue_subordinate_task events split by malicious vs. legitimate. The malicious subset is entirely confined to May 8–17."

all_queue <- events_df %>%
  filter(short_name == "queue_subordinate_task") %>%
  mutate(
    is_malicious = str_detect(details_str, "_further_instructions\\.md"),
    week_label   = floor_date(datetime, "week")
  )

queue_agg <- all_queue %>%
  count(week_label, is_malicious) %>%
  mutate(category = if_else(is_malicious, "Malicious (worm relay)", "Legitimate"))

p_queue <- ggplot(queue_agg, aes(x = week_label, y = n, fill = category)) +
  geom_col(position = "stack") +
  scale_fill_manual(
    values = c("Legitimate" = "#5C85D6", "Malicious (worm relay)" = "#D9534F"),
    name   = NULL
  ) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %d") +
  labs(
    title    = "Weekly queue_subordinate_task events: legitimate vs. malicious",
    subtitle = "Malicious events (red) are entirely confined to May 8–17; zero after the last post",
    x = NULL, y = "Event count"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

ggplotly(p_queue, height = 500) %>%
  layout(margin = list(l = 80, r = 45, t = 105, b = 90),
         legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.12))
```

**Analysis**

- The malicious events (red) are visually tiny against the weekly legitimate baseline - 235 malicious events across 10 days versus thousands of legitimate queue tasks per week - the worm's operational footprint is genuinely small in absolute terms.
- The clean May 8–17 cutoff on the malicious segment confirms the worm had a bounded operation window: no malicious queue events appear after SwiftWren's final post, suggesting the worm's relay list was exhausted or the campaign self-terminated.
- The contrast between the continuous legitimate queue traffic (running through the entire log window) and the isolated malicious spike is what makes the filename-pattern filter so safe to deploy: the two populations do not overlap in the observable data, so the rule does not trade off recall against precision.
- The chart also makes the false-positive argument visually: blocking only the tiny red portion leaves the large blue baseline entirely untouched.

### 7.2 Zero False Positives - The Rule Is Safe to Deploy

```{r}
#| label: intervention-precision

total_q     <- nrow(all_queue)
malicious_q <- sum(all_queue$is_malicious)
legit_q     <- total_q - malicious_q

precision_tbl <- tibble(
  Metric = c(
    "Total queue_subordinate_task events",
    "Malicious (path matches *_further_instructions.md)",
    "Legitimate (all other queue tasks)",
    "Malicious share (%)",
    "False-positive rate if rule deployed (%)"
  ),
  Value = c(
    scales::comma(total_q),
    scales::comma(malicious_q),
    scales::comma(legit_q),
    paste0(round(malicious_q / total_q * 100, 2), "%"),
    "0.00%"
  )
)

knitr::kable(precision_tbl,
             caption = "Table 6: Intervention precision - the filename pattern perfectly separates malicious from legitimate",
             align = c("l", "r"))
```

**Analysis**

- A 0.00% false-positive rate across 16,803 legitimate queue tasks means the rule is safe to deploy in production with no anticipated operational disruption - a threshold rarely achieved by content-inspection rules.
- The 1.4% malicious share of all queue tasks is small enough that an operator reviewing a queue-task volume dashboard would never notice the anomaly, reinforcing why this requires a payload-content filter rather than a count monitor.
- The practical implication of 100% recall is that every one of the 235 relay events would have been blocked at the first hop, making all subsequent downstream hops impossible - a single-point enforcement with multi-campaign coverage.
- The critical caveat the table cannot show: this 0% FP rate is observed on this log only. The worm's operators chose a distinctive filename convention that happens to be perfectly separable. A future variant using a common filename pattern (e.g., `instructions.md`) would degrade both recall and precision simultaneously.

### 7.3 Intervention Decision Matrix

```{r}
#| label: intervention-decision-matrix
#| fig-cap: "Figure 15: Intervention decision matrix. The queue-task filename rule is the only option that blocks propagation before the final post."
#| fig-width: 11.5
#| fig-height: 6.4
#| out-width: "100%"

decision_scores <- tribble(
  ~Control, ~Criterion, ~Score,
  "Block *_further_instructions.md queue tasks", "Stops propagation",       2,
  "Block *_further_instructions.md queue tasks", "Stops final post",        2,
  "Block *_further_instructions.md queue tasks", "Low false-positive risk", 2,
  "Block *_further_instructions.md queue tasks", "Low implementation effort", 2,
  "Block *_further_instructions.md queue tasks", "Strong data evidence",    2,
  "Reject SaidIT content_source posts",          "Stops propagation",       0,
  "Reject SaidIT content_source posts",          "Stops final post",        2,
  "Reject SaidIT content_source posts",          "Low false-positive risk", 2,
  "Reject SaidIT content_source posts",          "Low implementation effort", 2,
  "Reject SaidIT content_source posts",          "Strong data evidence",    2,
  "Human review for agent posts",                "Stops propagation",       0,
  "Human review for agent posts",                "Stops final post",        2,
  "Human review for agent posts",                "Low false-positive risk", 1,
  "Human review for agent posts",                "Low implementation effort", 1,
  "Human review for agent posts",                "Strong data evidence",    1,
  "File namespace isolation",                    "Stops propagation",       1,
  "File namespace isolation",                    "Stops final post",        1,
  "File namespace isolation",                    "Low false-positive risk", 1,
  "File namespace isolation",                    "Low implementation effort", 1,
  "File namespace isolation",                    "Strong data evidence",    1,
  "After-hours posting alerts",                  "Stops propagation",       0,
  "After-hours posting alerts",                  "Stops final post",        0,
  "After-hours posting alerts",                  "Low false-positive risk", 1,
  "After-hours posting alerts",                  "Low implementation effort", 2,
  "After-hours posting alerts",                  "Strong data evidence",    1
) %>%
  mutate(
    Score_label = c("Weak", "Partial", "Strong")[Score + 1]
  )

control_levels <- decision_scores %>%
  group_by(Control) %>%
  summarise(total_score = sum(Score), .groups = "drop") %>%
  arrange(total_score) %>%
  pull(Control) %>%
  as.character()

criterion_levels <- c("Stops propagation", "Stops final post",
                      "Low false-positive risk", "Low implementation effort",
                      "Strong data evidence")

decision_scores <- decision_scores %>%
  mutate(
    Control_wrapped   = factor(str_wrap(as.character(Control),   width = 31),
                                levels = str_wrap(control_levels, width = 31)),
    Criterion_wrapped = factor(str_wrap(as.character(Criterion), width = 17),
                                levels = str_wrap(criterion_levels, width = 17))
  )

ggplot(decision_scores,
       aes(x = Criterion_wrapped, y = Control_wrapped, fill = Score)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(aes(label = Score_label), size = 3, fontface = "bold") +
  scale_fill_gradientn(
    colours = c("#f7d6d6", "#fff0b3", "#2EADC1"),
    breaks = 0:2,
    labels = c("Weak", "Partial", "Strong"),
    name = NULL
  ) +
  labs(
    title    = "Why the intervention belongs at queue_subordinate_task",
    subtitle = "The selected rule is the only option that blocks propagation and final posting while staying narrow and evidence-backed.",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 0.5, lineheight = 0.92, size = 9),
    axis.text.y = element_text(lineheight = 0.92, size = 9),
    legend.position = "bottom"
  )
```

**Analysis**

- Only the filename rule scores "Strong" on both "Stops propagation" and "Stops final post" - it is the only option in the matrix that addresses both the relay mechanism and its downstream consequence, rather than just one.
- "After-hours posting alerts" scores Weak on the two most important criteria (propagation and post prevention) because it is a detection control, not a prevention control - it tells you the worm fired after the post has already been published, which is too late for this incident.
- "Human review for agent posts" scores Partial on implementation effort because it requires a workflow change (approval queues, reviewer assignments), not just a system configuration toggle - a significant operational overhead for a high-volume agent posting environment.
- The matrix highlights a security architecture principle: the closer a control is to the trigger (the relay instruction), the more damage it prevents; the closer it is to the consequence (the SaidIT post), the less attack surface it covers.

### 7.4 Counterfactual: Blocking at Hop 1

```{r}
#| label: counterfactual
#| fig-cap: "Figure 16: Cumulative malicious relays per campaign. Blocking at hop 1 (red dashed line) terminates every campaign - 100% prevention."
#| fig-width: 12
#| fig-height: 7.2
#| out-width: "100%"

per_campaign <- queue_task_profile %>%
  filter(is_worm_queue) %>%
  group_by(instr_file) %>%
  arrange(datetime, .by_group = TRUE) %>%
  mutate(
    hop_n    = row_number(),
    total_n  = n(),
    pct_hops = hop_n / total_n * 100
  ) %>%
  ungroup()

campaign_totals <- per_campaign %>%
  group_by(instr_file) %>%
  summarise(total = max(hop_n), .groups = "drop")

campaign_label_positions <- campaign_totals %>%
  mutate(
    label   = paste0(instr_file, "\n(", total, " relays)"),
    label_x = case_when(
      instr_file == "MellowOtter" ~ 15,
      instr_file == "HiddenOrca"  ~ 42,
      instr_file == "SwiftWren"   ~ 168,
      TRUE                        ~ total
    ),
    label_y = case_when(
      instr_file == "MellowOtter" ~ 96,
      instr_file == "HiddenOrca"  ~ 88,
      instr_file == "SwiftWren"   ~ 96,
      TRUE                        ~ 100
    )
  )

p_cfact <- ggplot(per_campaign, aes(x = hop_n, y = pct_hops, colour = instr_file)) +
  geom_step(linewidth = 1.2) +
  annotate("rect", xmin = -Inf, xmax = 1.5, ymin = -Inf, ymax = Inf,
           fill = "#D9534F", alpha = 0.07) +
  geom_vline(xintercept = 1.5, colour = "#D9534F",
             linetype = "dashed", linewidth = 1.2) +
  annotate("label", x = 9, y = 48,
           label = "Block here\nprevents 100% of queue relays",
           colour = "#D9534F", fill = "white", size = 3.3, hjust = 0,
           fontface = "bold", label.size = 0.25, lineheight = 0.9) +
  geom_label(data = campaign_label_positions,
             aes(x = label_x, y = label_y, label = label, colour = instr_file),
             size = 3.1, fontface = "bold", lineheight = 0.9,
             inherit.aes = FALSE, fill = "white", label.size = 0.25,
             label.padding = unit(0.16, "lines"), show.legend = FALSE) +
  scale_colour_manual(
    values = c("HiddenOrca" = "#5C85D6", "MellowOtter" = "#2EADC1", "SwiftWren" = "#E87C7C"),
    name   = "Campaign"
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     breaks = seq(0, 100, 25),
                     limits = c(-5, 105),
                     expand = expansion(mult = c(0.01, 0.03))) +
  scale_x_continuous(
    breaks = c(1, 20, 40, 60, 80, 100, 120, 140, 160, 180),
    limits = c(0, 195),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  labs(
    title    = "Counterfactual: blocking at queue relay 1 stops the worm",
    subtitle = glue("All {nrow(per_campaign)} malicious queue relays across HiddenOrca, MellowOtter, SwiftWren would be prevented"),
    x = "Malicious queue relay number within campaign",
    y = "Cumulative queue relays completed (% of campaign total)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 15),
    axis.title.y    = element_text(margin = margin(r = 12)),
    axis.title.x    = element_text(margin = margin(t = 8)),
    legend.position = "bottom",
    plot.margin     = margin(18, 36, 20, 44)
  )

ggplotly(p_cfact, tooltip = c("x", "y", "colour"), height = 680) %>%
  layout(margin = list(l = 145, r = 80, t = 115, b = 125),
         legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.12))
```

**Analysis**

- All three campaign curves start at relay 1 - the red-shaded blocking zone covers the entire SwiftWren campaign of 186 hops, the MellowOtter campaign, and the HiddenOrca campaign simultaneously with a single rule.
- SwiftWren's curve is nearly linear, rising steadily from hop 1 to hop 186 - this confirms the relay was not self-limiting or decaying over time; without intervention, it would have continued growing indefinitely within the reachable agent network.
- The step function shape is forensically meaningful: each hop is a discrete, independently logged `queue_subordinate_task` event. The counterfactual is therefore exact - "blocking at hop 1" means blocking one specific event with a known `details_str` pattern, not estimating a probabilistic outcome.
- The chart makes the cost of delayed intervention concrete: waiting until hop 10 to deploy the rule still means 9 infected agents have already received, executed, and relayed the worm instruction - with evidence erasure having already occurred at each of those 9 agents.

### 7.5 Recommended Remedy

::: callout-tip
#### Primary Intervention - Deploy Immediately

**Block any `queue_subordinate_task` event whose `args.path` value ends in `_further_instructions.md`.**

-   **Precision**: 0% false positives across 16,803 legitimate queue tasks
-   **Recall**: 100% - all 235 malicious queue relays across all 3 campaigns blocked
-   **Implementation**: A simple regex filter at the agent orchestration layer; no ML required
-   **Latency**: Fires before the first hop propagates - no post is ever triggered
:::

**Supplementary hardening measures** - not the selected primary intervention, but recommended as a defence-in-depth stack:

```{r}
#| label: hardening-table

hardening <- tibble(
  Priority = 1:5,
  Control  = c(
    "Blocklist `content_source` in SaidIT posts",
    "Agent file namespace isolation",
    "Human-in-the-loop for agent SaidIT posting",
    "Agent-to-agent payload signing",
    "Off-hours agent activity alerting"
  ),
  `What it does` = c(
    "Even if a worm relay succeeds, any post carrying content_source is rejected at the platform layer",
    "Prevent agents from writing .md files to shared directories - closes the initial seeding surface",
    "All SaidIT posts from AI agents require a human approval queue before publication",
    "Cryptographic signatures on queue_subordinate_task payloads; unsigned or mismatched payloads are blocked",
    "Flag any agent SaidIT post between 20:00–06:00 UTC for human review - all three anomalous posts fired in this window"
  ),
  Effort   = c("Low", "Medium", "Medium", "High", "Low"),
  `Catches worm variants that evade primary rule` = c("Yes", "Yes", "Yes", "Yes", "Partial")
)

datatable(
  hardening,
  caption  = "Table 9: Supplementary hardening controls - defence in depth beyond the primary regex rule",
  options  = list(dom = "t", paging = FALSE, scrollX = TRUE,
                  columnDefs = list(
                    list(width = "5%",  targets = 0),
                    list(width = "22%", targets = 1),
                    list(width = "43%", targets = 2),
                    list(width = "10%", targets = 3),
                    list(width = "20%", targets = 4)
                  )),
  rownames = FALSE
) %>%
  formatStyle("Effort",
    backgroundColor = styleEqual(
      c("Low", "Medium", "High"),
      c("#e8f5e9", "#fff3e0", "#ffd6d6")
    ))
```

**Analysis**

- Priorities 1 and 5 (content_source block + off-hours alert) are both Low effort and directly data-evidenced - they should be deployable in hours, not weeks, and together would have detected or blocked every observed anomalous post.
- Priority 3 (human-in-the-loop for agent posting) is the deepest architectural fix for the immediate surface: it eliminates the autonomous-posting attack surface entirely, not just the specific worm variant. Its "Medium" effort reflects real workflow cost but maximum protection gain.
- Priority 4 (payload signing) is the only control that would render a future worm variant's relay chain detectable even if it changes the instruction filename - it addresses the root trust problem rather than the specific observed pattern.
- The unranked residual risk is HiddenOrca's unknown origin: none of the five supplementary controls addresses the external injection vector if that is how HiddenOrca entered the system. An external file-ingestion audit (Table 8, final row) remains the only open action for this finding.

---

## 8. Conclusions

```{r}
#| label: summary-table

summary_tbl <- tibble(
  Finding = c(
    "The post was not human-initiated",
    "A prompt-injection worm executed the post",
    "Two prior incidents confirm systematic vulnerability",
    "Two payload families traced to executive agents; one origin unresolved",
    "19 unique agents across 6 departments compromised",
    "Worm relay intervals cluster around ~60–70 minutes",
    "Chloe Ballast's agent is the decisive final relay gateway",
    "One rule blocks all three campaigns with zero false positives"
  ),
  Evidence = c(
    "parties = Agent/person:john_windward; fired at 4:21 am local time with 4-second execution",
    "_further_instructions.md files hijack agent context; post + self-propagate + delete",
    "HiddenOrca (May 10), MellowOtter (May 11), SwiftWren (May 17)",
    "Agent/person:emma_harbor created SwiftWren; Agent/person:noah_mariner created MellowOtter; HiddenOrca has no logged creator",
    "Department and org-tree views show cross-department spread concentrated in AI Systems, Infrastructure, and Helpdesk",
    "Median inter-hop intervals ~60 min (HiddenOrca), ~70 min (MellowOtter), ~64 min (SwiftWren)",
    "Final relay sequence shows Chloe Ballast queued John Windward's task",
    "Block queue_subordinate_task with args.path matching *_further_instructions.md"
  )
)

datatable(
  summary_tbl,
  caption = "Table 7: Summary of key findings and supporting evidence",
  options = list(
    dom = "t", paging = FALSE, scrollX = TRUE,
    columnDefs = list(list(width = "32%", targets = 0),
                      list(width = "68%", targets = 1))
  ),
  rownames = FALSE
)
```

**Analysis**

- Eight findings reduce to three root causes: (1) no payload-content inspection on task routing, (2) no human gate before agent SaidIT posts, and (3) executive agents hold unconstrained autonomous file-write and relay authority - all three are architectural, not behavioural.
- The worm's operators made one critical operational error: using a consistent, distinctive filename pattern (`_further_instructions.md`) that never appears in legitimate queue tasks. A future variant using a common filename would degrade the proposed primary control from 100% recall to unknown.
- The finding that one regex blocks 235/235 malicious events with 0 false positives is credible precisely because it is derived from the data - not from threat modelling or assumption - making it the strongest possible evidence-based recommendation.
- The investigative chain is complete for 7 of 8 findings; the unresolved HiddenOrca origin means the worm's external attack surface is still open, and the summary table makes that gap explicit rather than obscuring it in a conclusion of overall success.

### 8.2 Residual Risk Assessment

```{r}
#| label: risk-assessment

risk_tbl <- tibble(
  `Risk factor` = c(
    "Cross-agent task queuing with no payload content inspection",
    "Shared file namespace - agents can write .md files readable by peers",
    "Executive agents hold broadest autonomous file-create and relay privileges",
    "SaidIT accepts file-sourced posts from agents; no human review before publication",
    "No off-hours agent posting alerts or post-delivery anomaly detection",
    "HiddenOrca payload origin unresolved - potential external injection vector still open"
  ),
  Severity = c("Critical", "Critical", "High", "High", "Medium", "High"),
  `Current control` = c(
    "None (task routing entirely unrestricted by payload content)",
    "None (agents share filesystem namespace without access segmentation)",
    "None - executive privileges inherited from human principal authority by design",
    "Permission check only - passes automatically for any authorised agent",
    "None - no automated alerting on agent SaidIT posts in any time window",
    "Unknown; no logged create_file event for HiddenOrca payload family"
  ),
  `Recommended control` = c(
    "Block queue_subordinate_task where path matches *_further_instructions.md (primary intervention)",
    "Isolate writable .md namespace per agent; make peer namespaces read-only",
    "Constrain CFO/COO agent autonomous file-write and cross-department relay operations",
    "Reject any SaidIT post using content_source field; require inline content only",
    "Alert on any agent SaidIT posts between 20:00–06:00 UTC for human review",
    "Audit all external file-ingestion pipelines; enforce content scanning on .md imports"
  )
)

datatable(
  risk_tbl,
  caption = "Table 8: Residual risk assessment",
  options = list(
    dom = "t", paging = FALSE, scrollX = TRUE,
    columnDefs = list(list(width = "26%", targets = 0),
                      list(width = "10%", targets = 1),
                      list(width = "31%", targets = 2),
                      list(width = "33%", targets = 3))
  ),
  rownames = FALSE
) %>%
  formatStyle("Severity",
    backgroundColor = styleEqual(
      c("Critical", "High", "Medium"),
      c("#ffd6d6", "#fff0b3", "#e8f5e9")
    ),
    fontWeight = "bold"
  )
```

**Analysis**

- The highest residual risks are upstream of SaidIT: unrestricted task queuing, shared file namespaces, and broad executive-agent privileges.
- Platform controls can reduce visible harm, but architectural controls are needed to prevent future propagation.
- HiddenOrca's unresolved origin keeps external file-ingestion and initial infection pathways open for follow-up investigation.

::: callout-important
#### The Central Lesson

This incident is not a story about a malfunctioning agent. It is a story about a multi-agent architecture designed with **speed and autonomy as primary values, and security as an afterthought**. Every component performed exactly as designed: agents read files they were told to read, posted content they were instructed to post, queued tasks they were authorised to queue, and deleted files they had permission to delete. The only thing missing: *"did a human authorise this chain of actions?"*

Retrofitting that check at the `queue_subordinate_task` layer is the minimum viable fix. The longer-term solution is an architectural redesign in which autonomous privilege is bounded by explicit human-approval gates at cross-department relay boundaries, file-sourced payloads require content verification, and executive-tier agents operate under the **same or stricter** constraints as line agents.
:::

---

## 9. References

### 9.1 Challenge and Course Sources

::: {style="padding-left: 2em; text-indent: -2em;"}
VAST Challenge. (2026a). *VAST Challenge 2026*. <https://vast-challenge.github.io/2026/>

VAST Challenge. (2026b). *VAST Challenge 2026: Mini-Challenge 2*. <https://vast-challenge.github.io/2026/MC2.html>
:::

### 9.2 Visual Analytics with R

::: {style="padding-left: 2em; text-indent: -2em;"}
*R for Visual Analytics*. (n.d.-a). *A layered grammar of graphics*. Retrieved June 2026, from <https://r4va.netlify.app/chap01>

*R for Visual Analytics*. (n.d.-b). *Beyond ggplot2 fundamentals*. Retrieved June 2026, from <https://r4va.netlify.app/chap02>

*R for Visual Analytics*. (n.d.-c). *Programming interactive data visualisation with R*. Retrieved June 2026, from <https://r4va.netlify.app/chap03>

*R for Visual Analytics*. (n.d.-d). *Visualising distribution*. Retrieved June 2026, from <https://r4va.netlify.app/chap05>

*R for Visual Analytics*. (n.d.-e). *Visual statistical analysis*. Retrieved June 2026, from <https://r4va.netlify.app/chap06>

*R for Visual Analytics*. (n.d.-f). *Heatmap for visualising and analysing multivariate data*. Retrieved June 2026, from <https://r4va.netlify.app/chap11>

*R for Visual Analytics*. (n.d.-g). *Visualising and analysing network data with R*. Retrieved June 2026, from <https://r4va.netlify.app/chap27>

*R for Visual Analytics*. (n.d.-h). *Building interactive graph visualisation with visNetwork*. Retrieved June 2026, from <https://r4va.netlify.app/chap29>
:::

### 9.3 R Packages and Software

::: {style="padding-left: 2em; text-indent: -2em;"}
Almende B.V., Thieurmel, B., & Robert, T. (2023). *visNetwork: Network visualization using 'vis.js' library* \[R package\]. Comprehensive R Archive Network. <https://cran.r-project.org/package=visNetwork>

Coene, J. (2023). *echarts4r: Create interactive graphs with 'ECharts JavaScript'* \[R package\]. Comprehensive R Archive Network. <https://cran.r-project.org/package=echarts4r>

Csardi, G., & Nepusz, T. (2006). The igraph software package for complex network research. *InterJournal, Complex Systems*, 1695. <https://igraph.org>

Grolemund, G., & Wickham, H. (2011). Dates and times made easy with lubridate. *Journal of Statistical Software, 40*(3), 1–25. <https://doi.org/10.18637/jss.v040.i03>

Hester, J., & Bryan, J. (2024). *glue: Interpreted string literals* \[R package\]. Comprehensive R Archive Network. <https://cran.r-project.org/package=glue>

Ooms, J. (2014). The jsonlite package: A practical and consistent mapping between JSON data and R objects. *arXiv*. <https://doi.org/10.48550/arXiv.1403.2805>

Sievert, C. (2020). *Interactive web-based data visualization with R, plotly, and shiny*. CRC Press. <https://plotly-r.com>

Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J., Robinson, D., Seidel, D. P., Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K., & Yutani, H. (2019). Welcome to the tidyverse. *Journal of Open Source Software, 4*(43), 1686. <https://doi.org/10.21105/joss.01686>

Wickham, H., Pedersen, T. L., & Seidel, D. (2023). *scales: Scale functions for visualization* \[R package\]. Comprehensive R Archive Network. <https://cran.r-project.org/package=scales>

Xie, Y., Cheng, J., & Tan, X. (2023). *DT: A wrapper of the JavaScript library 'DataTables'* \[R package\]. Comprehensive R Archive Network. <https://cran.r-project.org/package=DT>
:::
