---
title: "Methodology"
subtitle: "Analytical Approach for the TenantThread Forensics Investigation"
date: "2026-06-10"
date-modified: "last-modified"
execute:
  eval: false
  echo: true
  warning: false
  message: false
  freeze: true
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
---

## 1. Investigative Framework

This investigation applies a **forensic visual analytics workflow** to VAST Challenge 2026 MC2 - a 185,147-event log of a multi-agent AI system at Tenant Thread. The goal is not exploratory data analysis but bounded forensic reasoning: identify exactly what the data proves, distinguish it from what is merely suggested, and derive interventions that are empirically verifiable rather than heuristic.

The pipeline has six sequential stages, each validating its outputs before the next stage relies on them:

<div class="pipeline-flow">
  <div class="pipe-stage stage-1"><span class="pipe-num">1</span><span class="pipe-label">Data<br>Normalise</span></div>
  <span class="pipe-arrow">→</span>
  <div class="pipe-stage stage-2"><span class="pipe-num">2</span><span class="pipe-label">Feature<br>Engineer</span></div>
  <span class="pipe-arrow">→</span>
  <div class="pipe-stage stage-3"><span class="pipe-num">3</span><span class="pipe-label">Network<br>Build</span></div>
  <span class="pipe-arrow">→</span>
  <div class="pipe-stage stage-4"><span class="pipe-num">4</span><span class="pipe-label">Temporal<br>Analysis</span></div>
  <span class="pipe-arrow">→</span>
  <div class="pipe-stage stage-5"><span class="pipe-num">5</span><span class="pipe-label">Evidence<br>Evaluate</span></div>
  <span class="pipe-arrow">→</span>
  <div class="pipe-stage stage-6"><span class="pipe-num">6</span><span class="pipe-label">Intervention<br>Design</span></div>
</div>

---

## 2. Data Sources

Two JSON files from VAST Challenge 2026 MC2 drive every analysis on this site:

| File | Records | Key fields used |
|---|---|---|
| `MC2 data.json` | 185,147 events | `id`, `short_name`, `when`, `parties`, `details` |
| `org_chart.json` | Nodes + edges | `type`, `label`, `title`, `relation` |

Both files are read from `../VAST_Challenge_2026_MC2/` relative to the project root. No intermediate files are generated - every analytical object is derived fresh from these two sources.

---

## 3. Stage 1 - Data Normalisation

### Approach

The event log is a deeply nested JSON array where each event has a polymorphic `details` object - field names differ by event type. The key parsing decision is to use `simplifyVector = FALSE`, which preserves every event as an R list rather than coercing heterogeneous fields into a matrix that would silently drop values.

Each event is then flattened to five scalar fields via `map_dfr()`:

```{r}
events_df <- map_dfr(events_list, function(e) {
  tibble(
    id          = e$id          %||% NA_integer_,
    short_name  = e$short_name  %||% NA_character_,
    when        = e$when        %||% NA_real_,
    parties_str = paste(unlist(e$parties %||% list()), collapse = ";"),
    details_str = as.character(toJSON(e$details %||% list(), auto_unbox = TRUE))
  )
})
```

### Key design decisions

**`e$short_name`, not `e$event_type`** - The JSON field is `short_name`. Using `e$event_type` returns all-NA silently, producing no errors but wrong results downstream.

**`as.character(toJSON(...))`** - `toJSON()` returns class `json`, not `character`. Storing it without conversion causes `str_detect()` to fail with a class mismatch.

**`parties_str` vs `details_str`** - The *sending* agent is in `parties` (who issued the event); the *receiving* agent is in `details` (the task arguments). Using `parties_str` for both sender and receiver would produce sender == receiver for every relay event.

**Validation gate** - After flattening, an event-type frequency table confirms that `saidit_post`, `queue_subordinate_task`, `saidit_post_check`, `create_file`, and `delete_file` are all present before any analytical work begins. If any are absent, downstream analysis fails silently rather than loudly.

---

## 4. Stage 2 - Feature Engineering

### Approach

Six analytical objects are derived from the flat events table. Each targets one specific investigative question rather than creating a general-purpose data store.

| Object | Derived from | Answers |
|---|---|---|
| `chain_events` | `queue_subordinate_task` + `_further_instructions.md` path | Which agents relayed the worm, in what order? |
| `self_loops` | `chain_events` where sender == receiver | Zero-FP real-time detection signal |
| `post_events` | `saidit_post` / `post_saidit` + `content_source` field | Which posts are anomalous and why? |
| `perm_surface` | `saidit_post_check` | Which agents had SaidIT posting rights? |
| `person_org_lookup` | org chart edges (team and leadership paths) | Which department does each agent belong to? |
| `queue_task_profile` | All `queue_subordinate_task` events | What fraction are worm vs. legitimate? |

### Key design decisions

**Relay chain extraction** - The `_further_instructions.md` filename pattern is the discriminating signal. It appears in all worm relays and zero legitimate queue tasks:

```{r}
chain_events <- events_df %>%
  filter(
    short_name %in% c("queue_subordinate_task", "assign_agent_task"),
    str_detect(details_str, "_further_instructions\\.md")
  ) %>%
  mutate(
    sender   = str_remove(str_extract(parties_str, "Agent/person:[^;]+"),  "^Agent/"),
    receiver = str_remove(str_extract(details_str,  "Agent/person:[^\"]+"), "^Agent/")
  )
```

**Self-loops as a zero-FP signal** - An agent delegating a task to itself (`sender == receiver`) is structurally impossible in legitimate task flow. These 13 events appear only in worm relays and none in 16,803 legitimate queue tasks, making them a high-confidence real-time detection signal that requires no knowledge of filename or payload content.

**Two org lookup paths** - The org chart connects people to departments either via a `team:` intermediary or directly via a `department: → person:` edge (used for leadership). Without both paths, Executive Suite is silently dropped, producing an empty tab and missing the executives responsible for creating the worm payloads.

---

## 5. Stage 3 - Network Construction

### Approach

The worm relay chain is modelled as a **directed weighted graph** using `igraph`. Nodes are agents; directed edges represent relay events; edge weight is the count of repeated relays between the same pair. Betweenness centrality is computed on the normalised graph to assess whether any single agent acts as a structural chokepoint.

```{r}
ig <- graph_from_data_frame(
  chain_events %>%
    count(sender_clean, receiver_clean, name = "edge_weight") %>%
    rename(from = sender_clean, to = receiver_clean),
  directed = TRUE
)

btw <- betweenness(ig, normalized = TRUE)
```

### Why igraph for centrality, visNetwork for display

`igraph` is used for the centrality computation because it operates on the full graph structure without layout constraints. `visNetwork` is used for the interactive visualisation because it renders in HTML with physics-based layout and supports custom JavaScript hover events - neither is replaceable by the other for its task.

### Analytical implication of max betweenness ≈ 0.09

In a hub-and-spoke network, a hub scores above 0.5. A maximum of ~0.09 indicates a **distributed mesh with no structural chokepoint** - removing any single agent leaves viable relay paths intact. This result, derived at Stage 3, determines the intervention strategy at Stage 6: agent-level remediation is ineffective; the intervention must target the relay mechanism itself.

---

## 6. Stage 4 - Temporal Analysis

### Approach

Three temporal questions require different granularities:

| Question | Method | Why |
|---|---|---|
| What happened in the 4-second attack window? | Filter ±4s around confirmed post timestamp | Exact event ordering within a single second |
| How did the three campaigns differ in spread speed? | `ggplot2` density plot per campaign | Distribution shape - peaked vs. long-tailed - reveals structural differences invisible to summary statistics |
| Who created the payloads and when? | Filter `create_file` events for campaign filenames | Authoring fingerprint: 1-second spacing confirms automated, not manual, creation |

### Two-phase architecture

The campaign timeline reveals that relay infrastructure was seeded **17 hours before** the payload files were created. Detection strategies focused on the payload filename alone would miss the initial seeding phase entirely. This finding is only visible by examining relay events and file-creation events jointly on a shared timeline.

---

## 7. Stage 5 - Evidence Evaluation

### Approach

All findings are entered into a **forensic evidence register** that separates *observed* (directly readable from the log) from *inferred* (requires reasoning beyond the log). This prevents over-claiming in the conclusions section.

The central discriminating signal is the `content_source` field in the post `details` object:

- Appears in **all 3 anomalous posts**
- Appears in **zero of the 241 normal posts**

This binary separation is confirmed computationally rather than asserted. It establishes the reactive intervention (block posts containing `content_source`) with zero empirically observed false positives, and it is discovered at the feature engineering stage - before any network or temporal analysis begins.

### Separating observed from inferred

| Claim | Status |
|---|---|
| `queue_subordinate_task` with `_further_instructions.md` is the relay mechanism | **Observed** - 235 events in log |
| `content_source` perfectly separates anomalous from normal posts | **Observed** - 0/241 normal posts contain it |
| Emma Harbor created SwiftWren payload | **Observed** - `create_file` event logged |
| Noah Mariner created MellowOtter payload | **Observed** - `create_file` event logged |
| HiddenOrca payload origin | **Unresolved** - no `create_file` event in log |
| Exact content of deleted files | **Unresolvable** - log records deletion but not content |

---

## 8. Stage 6 - Intervention Design

### Approach

The intervention rule is derived directly from the feature engineering precision measure rather than designed independently:

> **Block `queue_subordinate_task` where `args.path` ends in `_further_instructions.md`**

This single regex rule blocks 235/235 malicious relays (100% recall) with 0 false positives across 16,803 legitimate queue events. The rule is proposed as Layer 1 of a three-layer defence:

| Layer | Control | Position in attack chain |
|---|---|---|
| **Layer 1** | Block `queue_subordinate_task` matching `_further_instructions.md` | Prevents relay from ever starting |
| **Layer 2** | Reject SaidIT posts with `content_source` field | Stops publication even if relay succeeds |
| **Layer 3** | Sandbox `.md` files - never parse as executable prompt context | Prevents context hijacking regardless of filename |

Layers 2 and 3 are proposed as belt-and-suspenders controls for worm variants that evade the filename pattern. Their precision is empirically verified for the observed campaigns; generalisation to unseen variants is architectural, not data-derived.

---

## 9. Visualisation Design Rationale

Each chart type was chosen for a specific geometric reason, not aesthetic preference:

| Visualisation | Used for | Geometric reason |
|---|---|---|
| echarts4r hierarchical tree | Org chart | Hierarchical containment maps naturally to a tree; interactive collapse prevents overplotting in large departments |
| ggplot2 area + `geom_vline` | Daily event volume | Temporal continuity emphasised; discrete anomaly markers do not obscure the baseline |
| visNetwork directed graph | Propagation network | Edge direction encodes information flow; node size by betweenness makes bridge topology visible without labelling every node |
| ggplot2 `geom_tile` heatmap | Dept-to-dept contamination; relay calendar | Matrix layout makes all pairwise flows legible simultaneously |
| ggplot2 `geom_segment` ladder | Relay sequence | Rank-ordered rows remove x-axis crowding; red terminal cell makes convergence visible |
| ggplot2 `geom_density` | Worm velocity | Distribution shape (peaked vs. long-tailed) reveals structure that histogram binning would mask |
| ggplot2 `geom_step` | Intervention counterfactual | Step function reflects the discrete "blocking a relay" event; shaded zone encodes the pre-rule region |
| DT `datatable` + `formatStyle` | Evidence tables | Colour-coded cells encode confidence and severity inline without requiring a separate legend |

---

## 10. Technology Stack

| Package | Role | Key pattern |
|---|---|---|
| `jsonlite` | JSON parsing | `fromJSON(..., simplifyVector = FALSE)` preserves nested lists |
| `tidyverse` | Wrangling + ggplot2 | `map_dfr()` for event flattening; `str_detect()` for field filtering |
| `lubridate` | Datetime handling | `as_datetime(when)` converts Unix seconds; `floor_date()` for period aggregation |
| `igraph` | Graph analytics | `graph_from_data_frame()` + `betweenness(normalized = TRUE)` |
| `visNetwork` | Interactive network | `visEvents()` for custom JS hover-highlight |
| `echarts4r` | Org chart tree | `e_list()` injects full ECharts JSON spec |
| `plotly` | Interactive charts | `ggplotly()` wraps ggplot2 for hover/zoom |
| `DT` | Interactive tables | `styleEqual()` for categorical colour mapping |
| `glue` | Dynamic text | Self-updating subtitles computed from runtime data |

---

## 11. Scope and Limitations

**What the log proves:**

- The worm relay mechanism, propagation path, and all three campaign timelines
- `content_source` as a binary discriminator between anomalous and normal posts
- The identity of payload authors for SwiftWren (Emma Harbor) and MellowOtter (Noah Mariner)
- The five-event attack sequence occurring within four seconds with no human approval step
- The filename-pattern block stops 100% of observed relays with 0% false positives in this log

**What the log does not prove:**

- Exact content of deleted payload files (deletion is logged; content is not)
- How `HiddenOrca_further_instructions.md` entered the system (no `create_file` event logged)
- Whether additional campaigns occurred outside the logging window
- What the agricultural vocabulary in `virus_check_in` events encodes
- Whether the worm was externally introduced or created internally under coercion or deception

::: callout-note
#### Reproducibility
All analysis runs against `../VAST_Challenge_2026_MC2/MC2 data.json` and `../VAST_Challenge_2026_MC2/org_chart.json` with no intermediate files. Re-running from scratch produces identical outputs.
:::
